---
title: 'Import - Lithic analysis Buhlen '
author: "Lisa Schunk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", 
  knit_root_dir = rprojroot::find_rstudio_root_file()) })
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script reads the xlsx file (database techno-typological analysis) generated with E4
and formats the data for a statistical analysis.  
The script will:

1. Reads in the original xlsx file 
2. Changes and sort the data  in order to do stats 
3. Saves the data as a new xlsx file and R object


```{r}
dir_in <-  "analysis/Buhlen/raw_data/"
dir_out <- "analysis/Buhlen/derived_data/"
```

Raw data must be located in "`r dir_in`".  
Formatted data will be saved in "`r dir_out`".
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(openxlsx)
library(readxl)
library(R.utils)
library(tools)
library(data.table)
library(chron)
library(dplyr)
```


---

# Get name, path and information of the file
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                      row.names = NULL)
```

The checksum (MD5 hashes) of the imported files are:  
```{r, echo = FALSE}
info_in
```


# Read in original xlsx-file
```{r}
imp_data <- read.xlsx(xlsxFile = data_file, sheet = 1, startRow = 1, colNames = TRUE,
                      rowNames = FALSE, skipEmptyRows = FALSE) 
#select the columns to keep 
keep_col <- c(1:2, 9:19, 21:35, 37:38, 40:43)
data_final <- imp_data[,keep_col]

data_final$length <- as.numeric(data_final$length)
data_final$width <- as.numeric(data_final$width)
data_final$thickness <- as.numeric(data_final$thickness)
data_final$weight <- as.numeric(data_final$weight)
data_final$perimeter.basis.back <- as.numeric(data_final$perimeter.basis.back)
data_final$perimeter.arch <- as.numeric(data_final$perimeter.arch)
data_final$perimeter.active.edge <- as.numeric(data_final$perimeter.active.edge)
data_final$perimeter.total <- as.numeric(data_final$perimeter.total)
data_final$thickness.back <- as.numeric(data_final$thickness.back)

str(data_final)

```

# Data analsysis - sorting  
## Dimension
```{r}
# keeps only columns relevant for dimensions and sorts them based on 
# their technological class
keep_col <- c(2, 4:5,19:21)
dimensions <- data_final[, keep_col] %>% arrange(technological.class)

KM_dimensions <- dimensions[2:131, ] %>% arrange(artefact.state)
KM.point_dimensions <- KM_dimensions[113:127, ]
KM.only_dimensions <- KM_dimensions[-(113:127), ]
KM.complete_dimensions <- KM_dimensions[1:111, ]
PS_dimensions <- dimensions[174:197, ] %>% arrange(artefact.state)
LSS_dimensions <- dimensions[132:173, ] %>% arrange(artefact.state)
S_dimensions <- dimensions[198:199, ] %>% arrange(artefact.state)


```

## Perimeter
```{r}
# keeps only columns relevant for perimeter measurements and sorts them 
# based on their technological class
keep_col <- c(2, 4:5,7,23:26)
perimeter <- data_final[, keep_col] %>% arrange(technological.class)

KM_perimeter <- perimeter[2:131, ] %>% arrange(artefact.state)
PS_perimeter <- perimeter[174:197, ] %>% arrange(artefact.state)

```

## Weight
```{r}
# keeps only columns relevant for weight measurements and sorts them based 
# on their technological class
keep_col <- c(2, 4:5, 22)
weight <- data_final[, keep_col] %>% arrange(technological.class)

KM_weight <- weight[2:131, ] %>% arrange(artefact.state)
PS_weight <- weight[174:197, ] %>% arrange(artefact.state)
LSS_weight <- weight[132:173, ] %>% arrange(artefact.state)
S_weight <- weight[198:199, ] %>% arrange(artefact.state)


```

## Raw material
```{r}
# keeps only columns relevant for raw material classification and sorts them 
# based on their technological class
keep_col <- c(2, 4:5, 3)
raw_material <- data_final[, keep_col] %>% arrange(technological.class)

KM_raw_material <- raw_material[2:131, ] %>% arrange(artefact.state)
PS_raw_material <- raw_material[174:197, ] %>% arrange(artefact.state)
LSS_raw_material <- raw_material[132:173, ] %>% arrange(artefact.state)
S_raw_material <- raw_material[198:199, ] %>% arrange(artefact.state)
 

```

## Cortex + blanks 
```{r}
# keeps only columns relevant for cortex and blank classification and sorts them 
# based on their technological class
keep_col <- c(2, 4:5, 6, 8:10)
cortex_blanks <- data_final[, keep_col] %>% arrange(technological.class)

KM_cortex_blanks <- cortex_blanks[2:131, ] %>% arrange(artefact.state)
PS_cortex_blanks <- cortex_blanks[174:197, ] %>% arrange(artefact.state)


```

## Back 
```{r}
# keeps only columns relevant for back modifications and sorts them based 
# on their technological class
keep_col <- c(2, 4:5, 11, 27)
back <- data_final[, keep_col] %>% arrange(technological.class)

KM_back <- back[2:131, ] %>% arrange(artefact.state)
PS_back <- back[174:197, ] %>% arrange(artefact.state)


```

## Edge retouch
```{r}
# keeps only columns relevant for edge retouch classification and sorts them 
# based on their technological class
keep_col <- c(2, 4:5, 12:13, 29)
edge_retouch <- data_final[, keep_col] %>% arrange(technological.class)

KM_edge_retouch <- edge_retouch[2:131, ] %>% arrange(artefact.state)
PS_edge_retouch <- edge_retouch[174:197, ] %>% arrange(artefact.state)
LSS_edge_retouch <- edge_retouch[132:173, ] %>% arrange(artefact.state)
S_edge_retouch <- edge_retouch[198:199, ] %>% arrange(artefact.state)


```

## Morpho type  
```{r}
# keeps only columns relevant for morpho type classification and sorts them 
# based on their technological class
keep_col <- c(2, 4:5, 7, 19:21)
morpho.type <- data_final[, keep_col] %>% arrange(technological.class)

KM_morpho.type <- morpho.type[2:131, ] %>% arrange(artefact.state)
PS_morpho.type <- morpho.type[174:197, ] %>% arrange(artefact.state)


```

## Application 'Pradnick method'  
```{r}
# keeps only columns relevant for 'morpho type 'Pradnick method' classification 
# and sorts them based on their technological class
keep_col <- c(2, 4:5, 15:16)
Pradnick.method <- data_final[, keep_col] %>% arrange(technological.class)

KM_Pradnick.method <- Pradnick.method[2:131, ] %>% arrange(artefact.state)
PS_Pradnick.method <- Pradnick.method[174:197, ] %>% arrange(artefact.state)


```

## Lateralisation
```{r}
# keeps only columns relevant for lateralisation and sorts them based on their
# technological class
keep_col <- c(2, 4:5, 18)
lateralisation <- data_final[, keep_col] %>% arrange(technological.class)

KM_lateralisation <- lateralisation[2:131, ] %>% arrange(artefact.state)
PS_lateralisation <- lateralisation[174:197, ] %>% arrange(artefact.state)


```

## Type lateral sharpening spall 
```{r}
# keeps only columns relevant for lateral sharpening spall classification and sorts 
# them based on their technological class
keep_col <- c(2, 4:5, 17:18)
lss_type <- data_final[, keep_col] %>% arrange(technological.class)

LSS_type <- lss_type[132:173, ] %>% arrange(artefact.state)


```


# Save data
## Format name of output file
```{r}
file_out <- "Buhlen_lithic_analysis"
```
The files will be saved as "`r paste0("~/", file_out, ".[ext]")`".


## Write to XLSX
```{r}
write.xlsx(list(data = data_final, dimensions = dimensions, KM_dimensions = KM_dimensions, KM.point_dimensions =
                  KM.point_dimensions, KM.only_dimensions = KM.only_dimensions,
                  KM.complete_dimensions = KM.complete_dimensions, PS_dimensions =
                  PS_dimensions, LSS_dimensions = LSS_dimensions, S_dimensions =
                  S_dimensions, KM_perimeter = KM_perimeter, PS_perimeter = PS_perimeter,
                  KM_weight = KM_weight, PS_weight = PS_weight,LSS_weight = LSS_weight,
                  S_weight = S_weight, KM_raw_material = KM_raw_material,PS_raw_material =
                  PS_raw_material, LSS_raw_material = LSS_raw_material, S_raw_material =
                  S_raw_material, KM_cortex_blanks = KM_cortex_blanks, PS_cortex_blanks =
                  PS_cortex_blanks, KM_back = KM_back, PS_back = PS_back, KM_edge_retouch =
                  KM_edge_retouch, PS_edge_retouch = PS_edge_retouch, LSS_edge_retouch =
                  LSS_edge_retouch, S_edge_retouch = S_edge_retouch, KM_morpho.type =
                  KM_morpho.type, PS_morpho.type = PS_morpho.type, KM_Pradnick.method =
                  KM_Pradnick.method, PS_Pradnick.method = PS_Pradnick.method,
                  KM_lateralisation = KM_lateralisation, PS_lateralisation =
                  PS_lateralisation, LSS_type = LSS_type),
                  file = paste0(dir_out, file_out, ".xlsx"))
```


## Save R object
```{r}
saveObject(data_final, file = paste0(dir_out, file_out, ".Rbin"))
```


## Show file information
```{r}
file_out <- c(paste0(dir_out, file_out, ".xlsx"), paste0(dir_out, file_out, ".Rbin"))
md5_out <- md5sum(file_out)
info_out <- data.frame(files = basename(names(md5_out)), checksum = md5_out, 
                       row.names = NULL)
```

The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---


END OF SCRIPT