---
title: 'Plots - Lithic analysis Balver Höhle '
author: "Lisa Schunk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", 
  knit_root_dir = rprojroot::find_rstudio_root_file()) })
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script reads the xlsx file (derived data) containing all the information gained through a lithic analysis.   
The script will:

1. Reads the xlsx file 
2. Plots all relevant variables in various combinations 
3. Saves the plot as PDFs


```{r}
dir_in <-  "analysis/Balve/derived_data/"
dir_out <- "analysis/Balve/plots/"
```

Raw data must be located in "`r dir_in`".  
Formatted data will be saved in "`r dir_out`".
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(openxlsx)
library(readxl)
library(R.utils)
library(tools)
library(chron)
library(ggplot2)
library(wesanderson)
library(dplyr)
library(ggsci)
```


---

# Get name, path and information of the file
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                      row.names = NULL)
```

The checksum (MD5 hashes) of the imported files are:  
```{r, echo = FALSE}
info_in
```


# Load data into R object
```{r}
imp_data <- read.xlsx(xlsxFile = data_file, sheet = 1, startRow = 1, colNames = TRUE,
                      rowNames = FALSE, skipEmptyRows = FALSE) 
```



# Data analsysis - plots  
## Histogram
### Histogram dimensions - Keilmesser 
```{r}
# Load data sheet Keilmesser 
KM_dim <- read.xlsx(xlsxFile = data_file, sheet = 3) 

# Keilmesser length 
# Calculates the mean value for the plot and ascribes the N value 
mean_length <- mean(KM_dim$length, na.rm = TRUE)
n <- doBy::summaryBy(length ~ artefact.state, data = KM_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Keilmesser length
KM.length <- ggplot(KM_dim, aes(x = length, fill = artefact.state)) + 
             geom_histogram(binwidth = 1) +
             labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
             theme_classic() +
             geom_vline(aes(xintercept = mean_length), linetype="dashed", size = 1) +
             geom_text(aes(y = mean_length, x = 70, label = round(mean_length, 1)), nudge_y = -48) +
             scale_fill_manual(values = wes_palette(n = 3, name = "Zissou1"), labels = tag) 

print(KM.length)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.length", ".pdf")
ggsave(filename = file_out, plot = KM.length, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Keilmesser width
# Calculates the mean value for the plot and ascribes the N value 
mean_width <- mean(KM_dim$width, na.rm = TRUE)
n <- doBy::summaryBy(width ~ artefact.state, data = KM_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Keilmesser width
KM.width <- ggplot(KM_dim, aes(x = width, fill = artefact.state)) + 
            geom_histogram(binwidth = 1) +
            labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
            theme_classic() +
            geom_vline(aes(xintercept = mean_width), linetype = "dashed", size = 1) +
            geom_text(aes(y = mean_length, x = 45, label = round(mean_width, 1)), nudge_y = -45) +
            scale_fill_manual(values = wes_palette(n = 3, name = "Zissou1"), labels = tag) 

print(KM.width) 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.width", ".pdf")
ggsave(filename = file_out, plot = KM.width, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Keilmesser thickness
# Calculates the mean value for the plot and ascribes the N value 
mean_thickness <- mean(KM_dim$thickness, na.rm = TRUE)
n <- doBy::summaryBy(thickness ~ artefact.state, data = KM_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Keilmesser thickness 
KM.thickness  <- ggplot(KM_dim, aes(y = thickness, fill = artefact.state)) + 
                 geom_histogram(binwidth = 1) +
                 labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
                 theme_classic() +
                 geom_hline(aes(yintercept = mean_thickness), linetype = "dashed", size = 1) +
                 geom_text(aes(y = mean_thickness, x = 20, label = round(mean_thickness, 1)), nudge_y = 1) +
                 scale_fill_manual(values = wes_palette(n = 3, name = "Zissou1"), labels = tag) 

print(KM.thickness) 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.thickness", ".pdf")
ggsave(filename = file_out, plot = KM.thickness, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")

# Keilmesser Weight
# Load data sheet Keilmesser weight
KM_weight <- read.xlsx(xlsxFile = data_file, sheet = 12) 

# Calculates the mean value for the plot and ascribes the N value 
mean_weight <- mean(KM_weight$weight, na.rm = TRUE)
n <- doBy::summaryBy(weight ~ artefact.state, data = KM_weight, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Keilmesser weight
KM.weight <- ggplot(KM_weight, aes(y = weight, fill = artefact.state)) + 
             geom_histogram(binwidth = 0.008) +
             labs(y = "weight [kg]", x = "N", title = "", fill = "artefact state", size = 12) +
             theme_classic() +
             geom_hline(aes(yintercept = mean_weight), linetype = "dashed", size = 1) +
             geom_text(aes(y = mean_weight, x = 20, label = round(mean_weight, 3)), nudge_y = 0.02) +
             scale_fill_manual(values = wes_palette(n = 3, name = "Zissou1"), labels = tag) 

print(KM.weight)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.weight", ".pdf")
ggsave(filename = file_out, plot = KM.weight, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")


# Keilmesser Back 
# Load data sheet Keilmesser thickness back 
KM_back <- read.xlsx(xlsxFile = data_file, sheet = 22) 
KM_back <- KM_back[-c(188, 190), ]


# Calculates the mean value for the plot and ascribes the N value 
mean_KM_back <- mean(KM_back$thickness.back, na.rm = TRUE)
n <- doBy::summaryBy(thickness.back ~ artefact.state, data = KM_back, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Keilmesser thickness back  
KM.back  <- ggplot(KM_back, aes(y = thickness.back, fill = artefact.state)) + 
                 geom_histogram(binwidth = 1) +
                 labs(x = "thickness [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
                 theme_classic() +
                 geom_hline(aes(yintercept = mean_KM_back), linetype = "dashed", size = 1) +
                 geom_text(aes(y = mean_KM_back, x = 16, label = round(mean_KM_back, 1)), nudge_y = -0.7) +
                 scale_fill_manual(values = wes_palette(n = 3, name = "Zissou1"), labels = tag) 

print(KM.back) 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.back", ".pdf")
ggsave(filename = file_out, plot = KM.back, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")

```


### Histogram dimensions - Pradnick scraper 
```{r}
# Load data sheet Pradnick scraper  
PS_dim <- read.xlsx(xlsxFile = data_file, sheet = 7) 

# Pradnick scraper length 
# Calculates the mean value for the plot and ascribes the N value 
mean_PS_length <- mean(PS_dim$length, na.rm = TRUE)
n <- doBy::summaryBy(length ~ artefact.state, data = PS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Pradnick scraper length
PS.length <- ggplot(PS_dim, aes(x = length, fill = artefact.state)) + 
             geom_histogram(binwidth = 1) +
             labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
             theme_classic() +
             geom_vline(aes(xintercept = mean_PS_length), linetype = "dashed", size = 1) +
             geom_text(aes(y = mean_PS_length, x = 55, label = round(mean_PS_length, 1)), nudge_y = -47) +
             scale_fill_manual(values = wes_palette(n = 1, name = "FantasticFox1"), labels = tag) 

print(PS.length)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.length", ".pdf")
ggsave(filename = file_out, plot = PS.length, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Pradnick scraper width 
# Calculates the mean value for the plot and ascribes the N value 
mean_PS_width <- mean(PS_dim$width, na.rm = TRUE)
n <- doBy::summaryBy(width ~ artefact.state, data = PS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Pradnick scraper width
PS.width <- ggplot(PS_dim, aes(x = width, fill = artefact.state)) + 
            geom_histogram(binwidth = 1) +
            labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
            theme_classic() +
            geom_vline(aes(xintercept = mean_PS_width), linetype="dashed", size = 1) +
            geom_text(aes(y = mean_PS_width, x = 30, label = round(mean_PS_width, 1)), nudge_y = -30) +
            scale_fill_manual(values = wes_palette(n = 1, name = "FantasticFox1"), labels = tag) 
 
print(PS.width)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.width", ".pdf")
ggsave(filename = file_out, plot = PS.width, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Pradnick scraper thickness
# Calculates the mean value for the plot and ascribes the N value 
mean_PS_thickness <- mean(PS_dim$thickness, na.rm = TRUE)
n <- doBy::summaryBy(thickness ~ artefact.state, data = PS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Pradnick scraper thickness 
PS.thickness <- ggplot(PS_dim, aes(y = thickness, fill = artefact.state)) + 
                geom_histogram(binwidth = 1) +
                labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
                theme_classic() +
                geom_hline(aes(yintercept = mean_PS_thickness), linetype = "dashed", size = 1) +
                geom_text(aes(y = mean_PS_thickness, x = 5.9, label = round(mean_PS_thickness, 1)), nudge_y = -1) +
                scale_fill_manual(values = wes_palette(n = 1 , name = "FantasticFox1"), labels = tag) 

print(PS.thickness)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.thickness", ".pdf")
ggsave(filename = file_out, plot = PS.thickness, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")


# Load data sheet Pradnick scraper weight
PS_weight <- read.xlsx(xlsxFile = data_file, sheet = 13) 

# Pradnick scraper weight
# Calculates the mean value for the plot and ascribes the N value 
mean_weight <- mean(PS_weight$weight, na.rm = TRUE)
n <- doBy::summaryBy(weight ~ artefact.state, data = PS_weight, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Pradnick scraper weight
PS.weight <- ggplot(PS_weight, aes(y = weight, fill = artefact.state)) + 
             geom_histogram(binwidth = 0.001) + 
             labs(y = "weight [kg]", x = "N", title = "", fill = "artefact state", size = 12) +
             theme_classic() +
             geom_hline(aes(yintercept = mean_weight), linetype="dashed", size = 1) +
             geom_text(aes(y = mean_weight, x= 2.9, label = round(mean_weight, 3)), nudge_y = -0.002) +
             scale_fill_manual(values = wes_palette(n =3, name = "Zissou1"), labels = tag) 
print(PS.weight)

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.weight", ".pdf")
ggsave(filename = file_out, plot = PS.weight, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")


# Back Pradnick scraper thickness
# Load data sheet Pradnick scraper thickness back 
PS_back <- read.xlsx(xlsxFile = data_file, sheet = 23) 

# Calculates the mean value for the plot and ascribes the N value 
mean_PS_back <- mean(PS_back$thickness.back, na.rm = TRUE)
n <- doBy::summaryBy(thickness.back ~ artefact.state, data = PS_back, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Pradnick scraper thickness back  
PS.back  <- ggplot(PS_back, aes(y = thickness.back, fill = artefact.state)) + 
                 geom_histogram(binwidth = 0.5) +
                 labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
                 theme_classic() +
                 geom_hline(aes(yintercept = mean_PS_back), linetype = "dashed", size = 1) +
                 geom_text(aes(y = mean_PS_back, x = 2.9, label = round(mean_PS_back, 1)), nudge_y = -0.8) +
                 scale_fill_manual(values = wes_palette(n = 3, name = "Zissou1"), labels = tag) 

print(PS.back) 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.back", ".pdf")
ggsave(filename = file_out, plot = PS.back, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")

```


### Histogram dimensions - Scraper 
```{r}
# Load data sheet Pradnick scraper  
S_dim <- read.xlsx(xlsxFile = data_file, sheet = 9) 

# Scraper length 
# Calculates the mean value for the plot and ascribes the N value 
mean_S_length <- mean(S_dim$length, na.rm = TRUE)
n <- doBy::summaryBy(length ~ artefact.state, data = S_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram scraper length
S.length <- ggplot(S_dim, aes(x = length, fill = artefact.state)) + 
            geom_histogram(binwidth = 1) +
            labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
            theme_classic() +
            geom_vline(aes(xintercept = mean_S_length), linetype = "dashed", size = 1) +
            geom_text(aes(y = mean_S_length, x = 60, label = round(mean_S_length, 2)), nudge_y = -64.1) +
            scale_fill_manual(values = wes_palette(n = 1, name = "Zissou1"), labels = tag)

print(S.length)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "S.length", ".pdf")
ggsave(filename = file_out, plot = S.length, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")



# Scraper width 
# Calculates the mean value for the plot and ascribes the N value 
mean_S_width <- mean(S_dim$width, na.rm = TRUE)
n <- doBy::summaryBy(width ~ artefact.state, data = S_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram scraper width
S.width <- ggplot(S_dim, aes(x = width, fill = artefact.state)) + 
           geom_histogram(binwidth = 1) +
           labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
           theme_classic() +
           geom_vline(aes(xintercept = mean_S_width), linetype = "dashed", size = 1) +
           geom_text(aes(y = mean_S_width, x = 35, label = round(mean_S_width, 1)), nudge_y = -36) +
           scale_fill_manual(values = wes_palette(n = 1, name = "Zissou1"), labels = tag)

print(S.width)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "S.width", ".pdf")
ggsave(filename = file_out, plot = S.width, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Scraper thickness
# Calculates the mean value for the plot and ascribes the N value 
mean_S_thickness <- mean(S_dim$thickness, na.rm = TRUE)
n <- doBy::summaryBy(thickness ~ artefact.state, data = S_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram scraper thickness 
S.thickness <- ggplot(S_dim, aes(y = thickness, fill = artefact.state)) + 
               geom_histogram(binwidth = 0.8) +
               labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
               theme_classic() +
               geom_hline(aes(yintercept = mean_S_thickness), linetype = "dashed", size = 1) +
               geom_text(aes(y = mean_S_thickness, x = 1.95, label = round(mean_S_thickness, 1)), nudge_y = -0.8) +
               scale_fill_manual(values = wes_palette(n = 1, name = "Zissou1"), labels = tag)

print(S.thickness)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "S.thickness", ".pdf")
ggsave(filename = file_out, plot = S.thickness, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")

# Load data sheet scraper weight
S_weight <- read.xlsx(xlsxFile = data_file, sheet = 15)

# Scraper weight
# Calculates the mean value for the plot and ascribes the N value 
mean_S_weight <- mean(S_weight$weight, na.rm = TRUE)
n <- doBy::summaryBy(weight ~ artefact.state, data = S_weight, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Pradnick scraper weight
S.weight <- ggplot(S_weight, aes(y = weight, fill = artefact.state)) + 
             geom_histogram(binwidth = 0.001) + 
             labs(y = "weight [kg]", x = "N", title = "", fill = "artefact state", size = 12) +
             theme_classic() +
             geom_hline(aes(yintercept = mean_S_weight), linetype="dashed", size = 1) +
             geom_text(aes(y = mean_S_weight, x= 1.05, label = round(mean_S_weight, 3)), nudge_y = -0.002) +
             scale_fill_manual(values = wes_palette(n =3, name = "Zissou1"), labels = tag) 

print(S.weight)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "S.weight", ".pdf")
ggsave(filename = file_out, plot = S.weight, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")

```


### Histogram dimension - Lateral sharpening spall
```{r}
# Load data sheet lateral sharpening spall  
LSS_dim <- read.xlsx(xlsxFile = data_file, sheet = 8) 

# Lateral sharpening spall length 
# Calculates the mean value for the plot and ascribes the N value 
mean_LSS_length <- mean(LSS_dim$length, na.rm = TRUE)
n <- doBy::summaryBy(length ~ artefact.state, data = LSS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))
  
# Histogram lateral sharpening spall length
LSS.length <- ggplot(LSS_dim, aes(x = length, fill = artefact.state)) + 
              geom_histogram(binwidth = 1) +
              labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
              theme_classic() +
              geom_vline(aes(xintercept = mean_LSS_length), linetype = "dashed", size = 1) +
              geom_text(aes(y = mean_LSS_length, x = 33, label = round(mean_LSS_length, 1)), nudge_y = -20) +
              scale_fill_manual(values = wes_palette(n = 3, name = "Darjeeling2"), labels = tag) 

print(LSS.length)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.length", ".pdf")
ggsave(filename = file_out, plot = LSS.length, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Lateral sharpening spall width 
# Calculates the mean value for the plot and ascribes the N value 
mean_LSS_width <- mean(LSS_dim$width, na.rm = TRUE)
n <- doBy::summaryBy(width ~ artefact.state, data = LSS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram lateral sharpening spall width
LSS.width <- ggplot(LSS_dim, aes(x = width, fill = artefact.state)) + 
             geom_histogram(binwidth = 1) +
             labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
             theme_classic() +
             geom_vline(aes(xintercept = mean_LSS_width), linetype = "dashed", size = 1) +
             geom_text(aes(y = mean_LSS_width, x = 19, label = round(mean_LSS_width, 1)), nudge_y = -7) +
             scale_fill_manual(values = wes_palette(n = 3, name = "Darjeeling2"), labels = tag) 

print(LSS.width)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.width", ".pdf")
ggsave(filename = file_out, plot = LSS.width, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Lateral sharpening spall thickness
# Calculates the mean value for the plot and ascribes the N value 
mean_LSS_thickness <- mean(LSS_dim$thickness, na.rm = TRUE)
n <- doBy::summaryBy(thickness ~ artefact.state, data = LSS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram lateral sharpening spall thickness 
LSS.thickness <- ggplot(LSS_dim, aes(y = thickness, fill = artefact.state)) + 
                 geom_histogram(binwidth = 0.8) +
                 labs(y = "thickness[mm]", x = "N", title = "", fill = "artefact state", size = 12) +
                 theme_classic() + 
                 geom_hline(aes(yintercept = mean_LSS_thickness), linetype = "dashed", size = 1) +
                 geom_text(aes(y = mean_LSS_thickness, x = 23.1, label = round(mean_LSS_thickness, 2)), nudge_y = -0.3) +
                 scale_fill_manual(values = wes_palette(n = 3, name = "Darjeeling2"), labels = tag) 

print(LSS.thickness)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.thickness", ".pdf")
ggsave(filename = file_out, plot = LSS.thickness, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")


# Load data sheet lateral sharpening spall weight
LSS_weight <- read.xlsx(xlsxFile = data_file, sheet = 14) 

# Scraper weight
# Calculates the mean value for the plot and ascribes the N value 
mean_LSS_weight <- mean(LSS_weight$weight, na.rm = TRUE)
n <- doBy::summaryBy(weight ~ artefact.state, data = LSS_weight, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Histogram Pradnick scraper weight
LSS.weight <- ggplot(LSS_weight, aes(y = weight, fill = artefact.state)) + 
             geom_histogram(binwidth = 0.001) + 
             labs(y = "weight [kg]", x = "N", title = "", fill = "artefact state", size = 12) +
             theme_classic() +
             geom_hline(aes(yintercept = mean_LSS_weight), linetype="dashed", size = 1) +
             geom_text(aes(y = mean_LSS_weight, x= 38, label = round(mean_LSS_weight, 3)), nudge_y = 0.0005) +
             scale_fill_manual(values = wes_palette(n =3, name = "Zissou1"), labels = tag) 

print(LSS.weight)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.weight", ".pdf")
ggsave(filename = file_out, plot = LSS.weight, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")

```

## Scatterplot 
### Length-width ratio 
```{r}
# Load data sheet Keilmesser 
KM_comp_dim <- read.xlsx(xlsxFile = data_file, sheet = 6) 

# Keilmesser length VS width 
# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ artefact.state, data = KM_comp_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Scatterplot Keilmesser (complete + point) length VS width 
KM.length_width <- ggplot(KM_comp_dim, aes(y = length, x = width, fill = artefact.state)) +
                   geom_point(size = 3, shape = 21) +
                   labs(x = "length [mm]", y = "width [mm]", title = "", fill = "artefact state", size = 12) +
                   xlim(10, 90) + ylim(0, 160) +
                   theme_classic() +
                   scale_fill_manual(values = wes_palette(n = 2, name = "FantasticFox1", type = "continuous"), labels = tag) 

print(KM.length_width)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.length_width", ".pdf")
ggsave(filename = file_out, plot = KM.length_width, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Pradnick scraper length VS width 
# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ artefact.state, data = PS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Scatterplot Pradnick scraper length VS width 
PS.length_width <- ggplot(PS_dim, aes(y = length, x = width, fill = artefact.state)) +
                   geom_point(size = 3, shape = 21) +
                   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "artefact state", size = 12) +
                   xlim(10, 70) + ylim(30, 80) +
                   theme_classic() +
                   scale_fill_manual(values = wes_palette(n =1,name = "FantasticFox1", type = "continuous"), labels = tag) 

print(PS.length_width)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.length_width", ".pdf")
ggsave(filename = file_out, plot = PS.length_width, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Lateral sharpening spall length VS width 
# Defines only the rows with complete LSS  
LSS.comp_dim <- LSS_dim[1:110,] 

# Lateral sharpening spall length VS width 
# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ artefact.state, data = LSS.comp_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Scatterplot lateral sharpening spall length VS width 
LSS.length_width <- ggplot(LSS.comp_dim, aes(y = length, x = width, fill = artefact.state)) +
                    geom_point(size = 3, shape = 21) +
                    labs(x = "length [mm]", y = "width [mm]", title = "", fill = "artefact state", size = 12) +
                    xlim(0, 45) + ylim(0, 65) +
                    theme_classic() +
                    scale_fill_manual(values = wes_palette(n = 1, name = "FantasticFox1", type = "continuous"), labels = tag) 

print(LSS.length_width)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.length_width", ".pdf")
ggsave(filename = file_out, plot = LSS.length_width, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")


# Keilmesser (complete) + Pradnick scraper length VS width
# Load data sheet dimensions  
dim <- read.xlsx(xlsxFile = data_file, sheet = 2) 
# Defines only the relevant rows 
KM.PS_dim <- dim[c(1:191, 309:335), ]  %>% arrange(artefact.state)
KM.PS_dim <- KM.PS_dim[1:185,]

# Keilmesser (complete) + Pradnick scraper length VS width 
# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ technological.class, data = KM.PS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))

# Scatterplot Keilmesser (complete) + Pradnick scraper length VS width
KM.PS.length_width <- ggplot(KM.PS_dim, aes(y = length, x = width, fill = technological.class)) +
                      geom_point(size = 3, shape = 21) +
                      labs(x = "length [mm]", y = "width [mm]", title = "",fill = "tool type", size = 12) +
                      xlim(0, 100) + ylim(0, 150) +
                      theme_classic() +
                      scale_fill_manual(values = wes_palette(n =3,name = "FantasticFox1", type = "continuous"), labels = tag) 

print(KM.PS.length_width)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.PS.length_width", ".pdf")
ggsave(filename = file_out, plot = KM.PS.length_width, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")
 
# Keilmesser (complete): length-width combined with morpho type
# Load data sheet Keilmesser morpho type 
KM_morpho.type <- read.xlsx(xlsxFile = data_file, sheet = 28)  

# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ morpho.type, data = KM_morpho.type, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]], ")"))


# Scatterplot Keilmesser (complete): length-width combined with morpho type 
KM.width_length_morpho <- ggplot(KM_morpho.type, aes(y = length, x = width, fill = morpho.type)) +
                      geom_point(size = 2, shape = 21) +
                      labs(x = "length [mm]", y = "width [mm]", title = "", fill = "morpho type", size = 12) +
                      xlim(0, 100) + ylim(0, 150) +
                      theme_classic() +
                      scale_fill_manual(values = wes_palette(n = 7,name = "FantasticFox1", type = "continuous"), labels = tag) 
                      
print(KM.width_length_morpho)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.width_length_morpho", ".pdf")
ggsave(filename = file_out, plot = KM.width_length_morpho, path = dir_out, device = "pdf", width = 170, height = 250, units = "mm")
```


## Barplot
### Morphotype 
```{r}
# Keilmesser morpho type 
# Load data sheet Keilmesser morpho type 
KM_morpho.type <- read.xlsx(xlsxFile = data_file, sheet = 28)  
# Defines only the rows with complete Keilmesser  
KM_morpho.type <- KM_morpho.type[1:158,]

# Barplot Keilmesser morpho type 
KM.morpho.type <- ggplot(data = KM_morpho.type) + aes(x = morpho.type, fill = morpho.type) + 
     geom_bar(stat = "count", width = 0.5) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "N") + 
     scale_fill_manual(values = wes_palette(n = 6,name = "Royal1", type = "continuous"))    
 
print(KM.morpho.type)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.morpho.type", ".pdf")
ggsave(filename = file_out, plot = KM.morpho.type, path = dir_out, device = "pdf", width = 170, height = 200, units = "mm")


# Pradnick scraper morpho type 
# Load data sheet Pradnick scraper morpho type 
PS_morpho.type <- read.xlsx(xlsxFile = data_file, sheet = 29)  

# Barplot Pradnick scraper morpho type 
PS.morpho.type <- ggplot(data = PS_morpho.type) + aes(x = morpho.type, fill = morpho.type) + 
     geom_bar(stat = "count", width = 0.36) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "N") + 
     scale_fill_manual(values = wes_palette(n = 4,name = "Royal1", type = "continuous"))    

print(PS.morpho.type)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.morpho.type", ".pdf")
ggsave(filename = file_out, plot = PS.morpho.type, path = dir_out, device = "pdf", width = 170, height = 200, units = "mm")

```


## Barplot
### Raw material 
```{r}
# Keilmesser raw material
# Load data sheet Keilmesser raw material
KM_raw_material <- read.xlsx(xlsxFile = data_file, sheet = 16)  

# Barplot Keilmesser raw material  
KM.raw_material <- ggplot(data = KM_raw_material) + aes(x = raw.material, fill = raw.material) + 
     geom_bar(stat = "count", width = 0.3) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "N") + 
     scale_x_discrete(labels=c("Baltic flint", "other", "Lydite", "N/A")) +
     scale_fill_manual(values = wes_palette(n = 4, name = "Darjeeling1", type = "continuous"))    
 
print(KM.raw_material)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.raw_material", ".pdf")
ggsave(filename = file_out, plot = KM.raw_material, path = dir_out, device = "pdf")


# Pradnick scraper raw material
# Load data sheet Pradnick scraper raw material
PS_raw_material <- read.xlsx(xlsxFile = data_file, sheet = 17)  

# Barplot Pradnick scraper raw material  
PS.raw_material <- ggplot(data = PS_raw_material) + aes(x = raw.material, fill = raw.material) + 
     geom_bar(stat = "count", width = 0.15) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "N") + 
     scale_x_discrete(labels=c("other", "Lydite")) +
     scale_fill_manual(values = wes_palette(n = 3, name = "Darjeeling1", type = "continuous"))    
 
print(PS.raw_material)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.raw_material", ".pdf")
ggsave(filename = file_out, plot = PS.raw_material, path = dir_out, device = "pdf")


# All tool types raw material
# Load data sheet all tool types raw material
all_raw_material <- read.xlsx(xlsxFile = data_file, sheet = 1)  

# Barplot Pradnick scraper raw material  
all.raw_material <- ggplot(data = all_raw_material) + aes(x = raw.material, fill = raw.material) + 
     geom_bar(stat = "count", width = 0.3) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "N") + 
     scale_x_discrete(labels=c("Baltic flint", "other", "Lydite", "N/A")) +
     scale_fill_manual(values = wes_palette(n = 4, name = "Darjeeling1", type = "continuous"))    
 
print(all.raw_material)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "all.raw_material", ".pdf")
ggsave(filename = file_out, plot = all.raw_material, path = dir_out, device = "pdf")
```


## Barplot
### Morphology back 
```{r}
# Keilmesser morphology back 
# Load data sheet Keilmesser raw material
KM_back <- read.xlsx(xlsxFile = data_file, sheet = 22)  

# Barplot Keilmesser raw material  
KM.back_morpho <- ggplot(data = KM_back) + aes(x = morphology.back, fill = morphology.back) + 
     geom_bar(stat = "count", width = 0.5) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "N") + 
     scale_x_discrete(labels=c("clear blunting", "cortex + partly retouched", "cortex/unworked", "N/A", "partly retouched", "retouched")) +
     scale_fill_manual(values = wes_palette(n = 7, name = "Darjeeling2", type = "continuous"))    
 
print(KM.back_morpho)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.back_morpho", ".pdf")
ggsave(filename = file_out, plot = KM.back_morpho, path = dir_out, device = "pdf",width = 250, height = 170, units = "mm")


# Pradnick scraper morphology back 
# Load data sheet Pradnick scraper raw material
PS_back <- read.xlsx(xlsxFile = data_file, sheet = 23)  

# Barplot Pradnick scraper raw material  
PS.back_morpho <- ggplot(data = PS_back) + aes(x = morphology.back, fill = morphology.back) + 
     geom_bar(stat = "count", width = 0.4) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "N") + 
     scale_x_discrete(labels=c("clear blunting", "cortex + partly retouched", "cortex/unworked", "N/A", "partly retouched", "retouched")) +
     scale_fill_manual(values = wes_palette(n = 7, name = "Darjeeling2", type = "continuous"))    
 
print(PS.back_morpho)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.back_morpho", ".pdf")
ggsave(filename = file_out, plot = PS.back_morpho, path = dir_out, device = "pdf",width = 250, height = 170, units = "mm")


# Keilmesser blanks
# Load data sheet Keilmesser blanks 
KM_cortex_blanks <- read.xlsx(xlsxFile = data_file, sheet = 20)  

# Barplot Keilmesser blanks 
KM.cortex_blanks <- ggplot(data = KM_cortex_blanks) + aes(x = blank, fill = cortex) + 
     geom_bar(stat = "count", width = 0.4) +
     theme_classic() +
     labs(x = " ", y = "N") + 
     scale_x_discrete(labels=c("core", "flake", "N/A")) +
     scale_fill_manual(values = wes_palette(n = 7, name = "Darjeeling2", type = "continuous"))    
 
print(KM.cortex_blanks)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.cortex_blanks", ".pdf")
ggsave(filename = file_out, plot = KM.cortex_blanks, path = dir_out, device = "pdf")


# Pradnick scraper blanks
# Load data sheet Pradnick scraper blanks 
PS_cortex_blanks <- read.xlsx(xlsxFile = data_file, sheet = 21)  

# Barplot Pradnick scraper blanks 
PS.cortex_blanks <- ggplot(data = PS_cortex_blanks) + aes(x = blank, fill = cortex) + 
     geom_bar(stat = "count", width = 0.3) +
     theme_classic() +
     labs(x = " ", y = "N") + 
     scale_x_discrete(labels=c("core", "flake", "N/A")) +
     scale_fill_manual(values = wes_palette(n = 7, name = "Darjeeling2", type = "continuous"))    
 
print(PS.cortex_blanks)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.cortex_blanks", ".pdf")
ggsave(filename = file_out, plot = PS.cortex_blanks, path = dir_out, device = "pdf")

```



## Barplot
### Pradnick method
```{r}
# Keilmesser application Pradnick method
# Load data sheet Keilmesser Pradnick method
KM_Pradnick.method <- read.xlsx(xlsxFile = data_file, sheet = 30)  

# Barplot Keilmesser Pradnick method 
KM.PM <- ggplot(data = KM_Pradnick.method) + aes(x = application.pradnick.method, fill = application.pradnick.method) + 
     geom_bar(stat = "count", width = 0.3) + 
     theme_classic() +
     labs(x = " ", y = "N") + 
     theme(legend.position = "none") + 
     scale_x_discrete(labels=c("no", "N/A", "yes")) +
     scale_fill_manual(values = wes_palette(n = 5, name = "GrandBudapest1", type = "continuous"))    
 
print(KM.PM)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.PM", ".pdf")
ggsave(filename = file_out, plot = KM.PM, path = dir_out, device = "pdf")

```


## Diverging barplot
### Lateralisation 
```{r}
# Keilmesser lateralisation
# Load data sheet Keilmesser lateralisation
KM_lateralisation <- read.xlsx(xlsxFile = data_file, sheet = 32)  
KM_lateralisation[grep("sin.", KM_lateralisation[["tool.lateralisation"]]), "value"] <- "-1"
KM_lateralisation[grep("dex.", KM_lateralisation[["tool.lateralisation"]]), "value"] <- "1"
# Defines only the rows with complete Keilmesser  
KM_lateralisation <- KM_lateralisation[1:186,]

# Barplot Keilmesser lateralisation
KM.lat <- ggplot(KM_lateralisation, aes(y = value, x = technological.class, fill = tool.lateralisation)) + 
          geom_bar(stat = "identity", width = 0.15) +
          theme_minimal() +
          labs(x = NULL, y = NULL) +
          theme(axis.text.x = element_blank(), axis.text.y = element_blank()) +
          scale_fill_manual(values = wes_palette(n = 5, name = "Darjeeling2", type = "continuous")) + 
          guides(fill = guide_legend(reverse=TRUE)) +
          labs(fill = "tool lateralisation") +
          coord_flip() 

print(KM.lat)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.lat", ".pdf")
ggsave(filename = file_out, plot = KM.lat, path = dir_out, device = "pdf")


# Pradnick scraper lateralisation
# Load data sheet Pradnick scraper lateralisation
PS_lateralisation <- read.xlsx(xlsxFile = data_file, sheet = 33)  
PS_lateralisation[grep("sin.", PS_lateralisation[["tool.lateralisation"]]), "value"] <- "-1"
PS_lateralisation[grep("dex.", PS_lateralisation[["tool.lateralisation"]]), "value"] <- "1"

# Barplot Pradnick scraper lateralisation
PS.lat <- ggplot(PS_lateralisation, aes(y = value, x = technological.class, fill = tool.lateralisation)) + 
          geom_bar(stat = "identity", width = 0.15) +
          theme_minimal() +
          labs(x = NULL, y = NULL) +
          theme(axis.text.x = element_blank(), axis.text.y = element_blank()) +
          scale_fill_manual(values = wes_palette(n = 5, name = "Darjeeling2", type = "continuous")) + 
          guides(fill = guide_legend(reverse=TRUE)) +
          labs(fill = "tool lateralisation") +
          coord_flip() 

print(PS.lat)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.lat", ".pdf")
ggsave(filename = file_out, plot = PS.lat, path = dir_out, device = "pdf")

```

## Barplot
### Barplot lateral resharpening spall type
```{r}
# Lateral resharpening spall type
# Load data sheet lateral resharpening spall type
LSS_type <- read.xlsx(xlsxFile = data_file, sheet = 34)  

# Barplot lateral resharpening spall type
LSS.type <- ggplot(data = LSS_type) + aes(x = type.lateral.sharpening.spall, fill = tool.lateralisation) + 
     geom_bar(stat = "count", width = 0.2) + 
     theme_classic() +
     labs(x = " ", y = "N") + 
     labs(fill = "tool lateralisation") +
     scale_x_discrete() +
     scale_fill_manual(values = wes_palette(n = 5, name = "GrandBudapest1", type = "continuous"))    
 
print(LSS.type)
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.type", ".pdf")
ggsave(filename = file_out, plot = LSS.type, path = dir_out, device = "pdf")

```



## Ternary plot
### Perimeter
```{r}
library(ggtern)
# Perimeter Keilmesser 
# Load data sheet Keilmesser perimeter
KM_perimeter <- read.xlsx(xlsxFile = data_file, sheet = 10) 
# Defines only the rows with complete Keilmesser  
KM_perimeter <- KM_perimeter[1:158,]

# Ternary diagram Keilmesser perimeter
KM.perimeter  <- ggtern(data = KM_perimeter, aes(x = perimeter.arch, y = perimeter.active.edge, z = perimeter.basis.back)) +
                 geom_point(aes(colour = morpho.type)) +
                 theme_bw() +
                 scale_colour_startrek() +
                 theme_hidetitles() +
                 theme_showarrows() +
                 xlab("arch")+ 
                 ylab("active edge")+
                 zlab("basis + back")+ 
                 labs(colour = "morpho type") +
                 tern_limits(labels=c(0, 20, 40, 60, 80, 100)) +
                 theme_rotate(degrees = 330)
                

print(KM.perimeter) 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.perimeter", ".pdf")
ggsave(filename = file_out, plot = KM.perimeter, path = dir_out, device = "pdf")


# Perimeter Pradnick scraper
# Load data sheet Pradnick scraper perimeter
PS_perimeter <- read.xlsx(xlsxFile = data_file, sheet = 11) 

# Ternary diagram Pradnick scraper perimeter
PS.perimeter  <- ggtern(data = PS_perimeter, aes(x = perimeter.arch, y = perimeter.active.edge, z = perimeter.basis.back)) +
                 geom_point(aes(colour = morpho.type)) +
                 theme_bw() +
                 scale_colour_startrek() +
                 theme_hidetitles() +
                 theme_showarrows() +
                 xlab("arch")+ 
                 ylab("active edge")+
                 zlab("basis + back")+ 
                 labs(colour = "morpho type") +
                 tern_limits(labels=c(0, 20, 40, 60, 80, 100)) +
                 theme_rotate(degrees = 330)
                

print(PS.perimeter) 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.perimeter", ".pdf")
ggsave(filename = file_out, plot = PS.perimeter, path = dir_out, device = "pdf")

```


---


# Show plot files information
```{r}
info_out <- list.files(path = dir_out, pattern = "\\.pdf$", full.names = TRUE) %>% 
            md5sum()
```


The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/Balve/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT







