---
title: 'Plots - Lithic analysis Balver Höhle '
author: "Lisa Schunk"
date: "`r Sys.Date()`"
output: word_document
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)

```


---

# Goal of the script
This script reads the xlsx file (derived data) containing all the information gained through a lithic analysis.   
The script will:

1. Read the xlsx file 
2. Plot all relevant variables in various combinations 
3. Save the plot as PDFs


```{r}
dir_in <-  "analysis/Balve/derived_data/"
dir_out <- "analysis/Balve/plots/"

```

Raw data must be located in "`r dir_in`".  
Formatted data will be saved in "`r dir_out`".
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(openxlsx)
library(readxl)
library(R.utils)
library(tools)
library(chron)
library(ggplot2)
library(wesanderson)
library(dplyr)
library(ggsci)

```


---

# Get name, path and information of the file
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                      row.names = NULL)

```

The checksum (MD5 hashes) of the imported files are:  
```{r, echo = FALSE}
info_in

```


# Load data into R object
```{r}
imp_data <- read.xlsx(xlsxFile = data_file, sheet = 1, startRow = 1, 
                      colNames = TRUE,
                      rowNames = FALSE, skipEmptyRows = FALSE) 

```



# Data analsysis - plots  
## Histogram
### Histogram dimensions - Keilmesser 
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Load data sheet Keilmesser 
KM_dim <- read.xlsx(xlsxFile = data_file, sheet = 3) 
KM_dim <- KM_dim [ , ] %>% arrange(artefact.state) 
KM.tip_dim <- KM_dim[-c(180:191), ]

# Keilmesser length 
# Calculates the mean value for the plot and ascribes the N value 
mean_length <- mean(KM.tip_dim$length, na.rm = TRUE)
n <- doBy::summaryBy(length ~ artefact.state, data = KM.tip_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                      ")"))

# Histogram Keilmesser length
KM.length <- ggplot(KM.tip_dim, aes(x = length, fill = artefact.state)) + 
             geom_histogram(binwidth = 1) +
             labs(x = "length [mm]", y = "n", title = "", 
             fill = "artefact state", size = 12) +
             theme_classic() +
             geom_vline(aes(xintercept = mean_length), linetype="dashed", 
             size = 1) +
             geom_text(aes(y = mean_length, x = 70, label = round(mean_length,  
                              1)), nudge_y = -48) +
             scale_fill_manual(values = wes_palette(n = 2, name =
             "FantasticFox1", type = "continuous"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.length", ".pdf")
ggsave(filename = file_out, plot = KM.length, path = dir_out, device = "pdf",
          width = 170, height = 250, units = "mm")


# Keilmesser width
# Calculates the mean value for the plot and ascribes the N value 
mean_width <- mean(KM.tip_dim$width, na.rm = TRUE)
n <- doBy::summaryBy(width ~ artefact.state, data = KM.tip_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                        ")"))

# Histogram Keilmesser width
KM.width <- ggplot(KM.tip_dim, aes(x = width, fill = artefact.state)) + 
            geom_histogram(binwidth = 1) +
            labs(x = "width [mm]", y = "n", title = "", fill = "artefact state",
            size = 12) +
            theme_classic() +
            geom_vline(aes(xintercept = mean_width), linetype = "dashed", 
            size = 1) +
            geom_text(aes(y = mean_length, x = 45, label = round(mean_width, 1)),
            nudge_y = -45) +
            scale_fill_manual(values = wes_palette(n = 2, name = "FantasticFox1",
            type = "continuous"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.width", ".pdf")
ggsave(filename = file_out, plot = KM.width, path = dir_out, device = "pdf",
          width = 170, height = 250, units = "mm")


# Keilmesser thickness
# Calculates the mean value for the plot and ascribes the N value 
mean_thickness <- mean(KM.tip_dim$thickness, na.rm = TRUE)
n <- doBy::summaryBy(thickness ~ artefact.state, data = KM.tip_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                                     ")"))

# Histogram Keilmesser thickness 
KM.thickness  <- ggplot(KM.tip_dim, aes(y = thickness, fill = artefact.state)) + 
                 geom_histogram(binwidth = 1) +
                 labs(y = "thickness [mm]", x = "n", title = "", 
                 fill = "artefact state", size = 12) +
                 theme_classic() +
                 geom_hline(aes(yintercept = mean_thickness), 
                 linetype = "dashed", size = 1) +
                 geom_text(aes(y = mean_thickness, x = 20, 
                 label = round(mean_thickness, 1)), nudge_y = 1) +
                 scale_fill_manual(values = wes_palette(n = 2, 
                 name = "FantasticFox1", type = "continuous"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.thickness", ".pdf")
ggsave(filename = file_out, plot = KM.thickness, path = dir_out, device = "pdf",
          width = 250, height = 170, units = "mm")


# Keilmesser Back 
# Load data sheet Keilmesser thickness back 
KM_back <- read.xlsx(xlsxFile = data_file, sheet = 22) 
KM_back <- KM_back[-c(159:191), ]

# Calculates the mean value for the plot and ascribes the N value 
mean_KM_back <- mean(KM_back$thickness.back, na.rm = TRUE)
n <- doBy::summaryBy(thickness.back ~ artefact.state, data = KM_back, 
                     FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], 
                                         " (n = ", n[[2]], ")"))

# Histogram Keilmesser thickness back  
KM.back  <- ggplot(KM_back, aes(y = thickness.back, fill = artefact.state)) + 
                 geom_histogram(binwidth = 1) +
                 labs(x = "thickness [mm]", y = "n", title = "", 
                 fill = "artefact state", size = 12) +
                 theme_classic() +
                 geom_hline(aes(yintercept = mean_KM_back), linetype = "dashed",
                 size = 1) +
                 geom_text(aes(y = mean_KM_back, x = 16, 
                 label = round(mean_KM_back, 1)), nudge_y = -0.7) +
                 scale_fill_manual(values = wes_palette(n = 2, 
                 name = "FantasticFox1", type = "continuous"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.back", ".pdf")
ggsave(filename = file_out, plot = KM.back, path = dir_out, device = "pdf", 
          width = 250, height = 170, units = "mm")

```


### Histogram dimensions - Pradnik scraper 
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Load data sheet Pradnik scraper  
PS_dim <- read.xlsx(xlsxFile = data_file, sheet = 7) 

# Pradnik scraper length 
# Calculates the mean value for the plot and ascribes the N value 
mean_PS_length <- mean(PS_dim$length, na.rm = TRUE)
n <- doBy::summaryBy(length ~ artefact.state, data = PS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]],
                                         " (n = ", n[[2]], ")"))

# Histogram Pradnik scraper length
PS.length <- ggplot(PS_dim, aes(x = length, fill = artefact.state)) + 
             geom_histogram(binwidth = 1) +
             labs(x = "length [mm]", y = "n", title = "", 
             fill = "artefact state", size = 12) +
             theme_classic() +
             geom_vline(aes(xintercept = mean_PS_length), linetype = "dashed",
             size = 1) +
             geom_text(aes(y = mean_PS_length, x = 55, 
             label = round(mean_PS_length, 1)), nudge_y = -47) +
             scale_fill_manual(values = wes_palette(n = 1, name = "Chevalier1"),
             labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.length", ".pdf")
ggsave(filename = file_out, plot = PS.length, path = dir_out, device = "pdf",
          width = 170, height = 250, units = "mm")


# Pradnik scraper width 
# Calculates the mean value for the plot and ascribes the N value 
mean_PS_width <- mean(PS_dim$width, na.rm = TRUE)
n <- doBy::summaryBy(width ~ artefact.state, data = PS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]],
                                         " (n = ", n[[2]], ")"))

# Histogram Pradnik scraper width
PS.width <- ggplot(PS_dim, aes(x = width, fill = artefact.state)) + 
            geom_histogram(binwidth = 1) +
            labs(x = "width [mm]", y = "n", title = "", fill = "artefact state",
            size = 12) +
            theme_classic() +
            geom_vline(aes(xintercept = mean_PS_width), linetype="dashed",
            size = 1) +
            geom_text(aes(y = mean_PS_width, x = 30, label = round(mean_PS_width,
            1)), nudge_y = -30) +
            scale_fill_manual(values = wes_palette(n = 1, name = "Chevalier1"), labels = tag) 
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.width", ".pdf")
ggsave(filename = file_out, plot = PS.width, path = dir_out, device = "pdf",
          width = 170, height = 250, units = "mm")


# Pradnik scraper thickness
# Calculates the mean value for the plot and ascribes the N value 
mean_PS_thickness <- mean(PS_dim$thickness, na.rm = TRUE)
n <- doBy::summaryBy(thickness ~ artefact.state, data = PS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], 
                                         " (n = ", n[[2]], ")"))

# Histogram Pradnik scraper thickness 
PS.thickness <- ggplot(PS_dim, aes(y = thickness, fill = artefact.state)) + 
                geom_histogram(binwidth = 1) +
                labs(y = "thickness [mm]", x = "n", title = "", 
                fill = "artefact state", size = 12) +
                theme_classic() +
                geom_hline(aes(yintercept = mean_PS_thickness), 
                linetype = "dashed", size = 1) +
                geom_text(aes(y = mean_PS_thickness, x = 5.9, 
                label = round(mean_PS_thickness, 1)), nudge_y = -1) +
                scale_fill_manual(values = wes_palette(n = 1, name =
                "Chevalier1"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.thickness", ".pdf")
ggsave(filename = file_out, plot = PS.thickness, path = dir_out, device = "pdf",
          width = 250, height = 170, units = "mm")


# Back Pradnik scraper thickness
# Load data sheet Pradnik scraper thickness back 
PS_back <- read.xlsx(xlsxFile = data_file, sheet = 23) 

# Calculates the mean value for the plot and ascribes the N value 
mean_PS_back <- mean(PS_back$thickness.back, na.rm = TRUE)
n <- doBy::summaryBy(thickness.back ~ artefact.state, data = PS_back, 
                     FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                                     ")"))

# Histogram Pradnik scraper thickness back  
PS.back  <- ggplot(PS_back, aes(y = thickness.back, fill = artefact.state)) + 
                 geom_histogram(binwidth = 0.8) +
                 labs(y = "thickness [mm]", x = "n", title = "", 
                 fill = "artefact state", size = 12) +
                 theme_classic() +
                 geom_hline(aes(yintercept = mean_PS_back), linetype = "dashed",
                 size = 1) +
                 geom_text(aes(y = mean_PS_back, x = 2.9, 
                 label = round(mean_PS_back, 1)), nudge_y = -0.8) +
                 scale_fill_manual(values = wes_palette(n = 1, name =
                 "Chevalier1"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.back", ".pdf")
ggsave(filename = file_out, plot = PS.back, path = dir_out, device = "pdf", 
          width = 250, height = 170, units = "mm")

```



### Histogram dimension - Lateral sharpening spall
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Load data sheet lateral sharpening spall  
LSS_dim <- read.xlsx(xlsxFile = data_file, sheet = 8) 
LSS_dim <- LSS_dim [-c(111:117), ]

# Lateral sharpening spall length 
# Calculates the mean value for the plot and ascribes the N value 
mean_LSS_length <- mean(LSS_dim$length, na.rm = TRUE)
n <- doBy::summaryBy(length ~ artefact.state, data = LSS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (N = ", n[[2]],
                                                     ")"))
# Histogram lateral sharpening spall length
LSS.length <- ggplot(LSS_dim, aes(x = length, fill = artefact.state)) + 
              geom_histogram(binwidth = 1) +
              labs(x = "length [mm]", y = "N", title = "", 
              fill = "artefact state", size = 12) +
              theme_classic() +
              geom_vline(aes(xintercept = mean_LSS_length), linetype = "dashed",
              size = 1) +
              geom_text(aes(y = mean_LSS_length, x = 33, 
              label = round(mean_LSS_length, 1)), nudge_y = -20) +
              scale_fill_manual(values = wes_palette(n = 1, name = "Royal1"),
              labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.length", ".pdf")
ggsave(filename = file_out, plot = LSS.length, path = dir_out, device = "pdf",
          width = 170, height = 250, units = "mm")


# Lateral sharpening spall width 
# Calculates the mean value for the plot and ascribes the N value 
mean_LSS_width <- mean(LSS_dim$width, na.rm = TRUE)
n <- doBy::summaryBy(width ~ artefact.state, data = LSS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                                     ")"))

# Histogram lateral sharpening spall width
LSS.width <- ggplot(LSS_dim, aes(x = width, fill = artefact.state)) + 
             geom_histogram(binwidth = 1) +
             labs(x = "width [mm]", y = "n", title = "", fill = "artefact state",
             size = 12) +
             theme_classic() +
             geom_vline(aes(xintercept = mean_LSS_width), linetype = "dashed",
             size = 1) +
             geom_text(aes(y = mean_LSS_width, x = 19, 
             label = round(mean_LSS_width, 1)), nudge_y = -7) +
             scale_fill_manual(values = wes_palette(n = 1, name = "Royal1"),
                               labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.width", ".pdf")
ggsave(filename = file_out, plot = LSS.width, path = dir_out, device = "pdf",
          width = 170, height = 250, units = "mm")


# Lateral sharpening spall thickness
# Calculates the mean value for the plot and ascribes the N value 
mean_LSS_thickness <- mean(LSS_dim$thickness, na.rm = TRUE)
n <- doBy::summaryBy(thickness ~ artefact.state, data = LSS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                                     ")"))

# Histogram lateral sharpening spall thickness 
LSS.thickness <- ggplot(LSS_dim, aes(y = thickness, fill = artefact.state)) + 
                 geom_histogram(binwidth = 0.8) +
                 labs(y = "thickness[mm]", x = "n", title = "", 
                 fill = "artefact state", size = 12) +
                 theme_classic() + 
                 geom_hline(aes(yintercept = mean_LSS_thickness), 
                 linetype = "dashed", size = 1) +
                 geom_text(aes(y = mean_LSS_thickness, x = 23.1, 
                 label = round(mean_LSS_thickness, 2)), nudge_y = -0.3) +
                 scale_fill_manual(values = wes_palette(n = 1, name = "Royal1"),
                 labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.thickness",
                   ".pdf")
ggsave(filename = file_out, plot = LSS.thickness, path = dir_out, 
         device = "pdf", width = 250, height = 170, units = "mm")

```

## Scatterplot 
### Length-width ratio 
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Load data sheet Keilmesser 
KM_comp_dim <- read.xlsx(xlsxFile = data_file, sheet = 6) 

# Keilmesser length VS width 
# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ artefact.state, data = KM_comp_dim, 
                     FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                                     ")"))

# Scatterplot Keilmesser (complete + point) length VS width 
KM.length_width <- ggplot(KM_comp_dim, aes(y = length, x = width, 
                                        fill = artefact.state)) +
                   geom_point(size = 3, shape = 21) +
                   labs(y = "length [mm]", x = "width [mm]", title = "", 
                   fill = "artefact state", size = 12) +
                   xlim(0, 90) + ylim(0, 160) +
                   theme_classic() +
                   scale_fill_manual(values = wes_palette(n = 2, 
                   name = "FantasticFox1", type = "continuous"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.length_width",
                   ".pdf")
ggsave(filename = file_out, plot = KM.length_width, path = dir_out, 
         device = "pdf", width = 170, height = 250, units = "mm")


# Pradnik scraper length VS width 
# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ artefact.state, data = PS_dim, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                                     ")"))

# Scatterplot Pradnik scraper length VS width 
PS.length_width <- ggplot(PS_dim, aes(y = length, x = width, 
                                   fill = artefact.state)) +
                   geom_point(size = 3, shape = 21) +
                   labs(y = "length [mm]", x = "width [mm]", title = "", 
                   fill = "artefact state", size = 12) +
                   xlim(0, 70) + ylim(0, 80) +
                   theme_classic() +
                   scale_fill_manual(values = wes_palette(n = 6, 
                   name = "Chevalier1", type = "continuous"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.length_width",
                   ".pdf")
ggsave(filename = file_out, plot = PS.length_width, path = dir_out, 
       device = "pdf", width = 170, height = 250, units = "mm")


# Lateral sharpening spall length VS width 
# Defines only the rows with complete LSS  
LSS.comp_dim <- LSS_dim[1:110,] 

# Lateral sharpening spall length VS width 
# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ artefact.state, data = LSS.comp_dim, 
                     FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                             ")"))

# Scatterplot lateral sharpening spall length VS width 
LSS.length_width <- ggplot(LSS.comp_dim, aes(y = length, x = width, 
                                          fill = artefact.state)) +
                    geom_point(size = 3, shape = 21) +
                    labs(y = "length [mm]", x = "width [mm]", title = "", 
                    fill = "artefact state", size = 12) +
                    xlim(0, 45) + ylim(0, 65) +
                    theme_classic() +
                    scale_fill_manual(values = wes_palette(n = 2, 
                    name = "Royal1", type = "continuous"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.length_width",
                   ".pdf")
ggsave(filename = file_out, plot = LSS.length_width, path = dir_out, 
         device = "pdf", width = 170, height = 250, units = "mm")


# Keilmesser (complete) + Pradnik scraper length VS width
# Load data sheet dimensions  
dim <- read.xlsx(xlsxFile = data_file, sheet = 2) 
# Defines only the relevant rows 
KM.PS_dim <- dim[c(1:191, 309:335), ]  %>% arrange(artefact.state)
KM.PS_dim <- KM.PS_dim[1:185,]

# Keilmesser (complete) + Pradnik scraper length VS width 
# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ technological.class, data = KM.PS_dim, 
                     FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                                     ")"))

# Scatterplot Keilmesser (complete) + Pradnik scraper length VS width
KM.PS.length_width <- ggplot(KM.PS_dim, aes(y = length, x = width, 
                                         fill = technological.class)) +
                      geom_point(size = 3, shape = 21) +
                      labs(y = "length [mm]", x = "width [mm]", title = "",
                      fill = "artefact category", size = 12) +
                      xlim(0, 100) + ylim(0, 150) +
                      theme_classic() +
                      scale_fill_manual(values = wes_palette(n = 3, name =
                      "GrandBudapest1", type = "continuous"), labels = tag) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.PS.length_width",
                   ".pdf")
ggsave(filename = file_out, plot = KM.PS.length_width, path = dir_out, 
       device = "pdf", width = 170, height = 250, units = "mm")
 
# Keilmesser (complete): length-width combined with morpho type
# Load data sheet Keilmesser morpho type 
KM_morpho.type <- read.xlsx(xlsxFile = data_file, sheet = 28)  

# Ascribes the N value 
n <- doBy::summaryBy(length + width ~ morpho.type, data = KM_morpho.type, 
                     FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                                     ")"))


# Scatterplot Keilmesser (complete): length-width combined with morpho type 
KM.width_length_morpho <- ggplot(KM_morpho.type, aes(y = length, x = width, 
                                                  fill = morpho.type)) +
                      geom_point(size = 2, shape = 21) +
                      labs(y = "length [mm]", x = "width [mm]", title = "", 
                      fill = "Keilmesser shape", size = 12) +
                      xlim(0, 100) + ylim(0, 150) +
                      theme_classic() +
                      scale_fill_manual(values = wes_palette(n = 7,name =
                      "FantasticFox1", type = "continuous"), labels = tag) 
                      
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.width_length_morpho", ".pdf")
ggsave(filename = file_out, plot = KM.width_length_morpho, path = dir_out, 
         device = "pdf", width = 170, height = 250, units = "mm")

```


## Barplot
### Morphotype 
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Keilmesser morpho type 
# Load data sheet Keilmesser morpho type 
KM_morpho.type <- read.xlsx(xlsxFile = data_file, sheet = 28)  
# Defines only the rows with complete Keilmesser  
KM_morpho.type <- KM_morpho.type[1:158,]

# Barplot Keilmesser morpho type 
KM.morpho.type <- ggplot(data = KM_morpho.type) + aes(x = morpho.type, 
                                                   fill = morpho.type) + 
     geom_bar(stat = "count", width = 0.5, fill = c("#899DA4", "#A46F65",
                    "#BF4226", "#D76848", "#EBB99A", "#E8B37B")) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "n") 
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.morpho.type",
                   ".pdf")
ggsave(filename = file_out, plot = KM.morpho.type, path = dir_out, 
         device = "pdf", width = 170, height = 200, units = "mm")

```


## Barplot
### Edge retouch  
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Keilmesser edge retouch 
# Load data sheet Keilmesser edge retouch 
KM_edge <- read.xlsx(xlsxFile = data_file, sheet = 24)  
# Defines only the rows with complete Keilmesser and Keilmesser tips   
KM_edge <- KM_edge[-c(180:191),]

# Barplot Keilmesser edge retouch  
KM.edge <- ggplot(data = KM_edge) + aes(x = retouch.type.edge, 
                                     fill = retouch.type.edge) + 
     geom_bar(stat = "count", width = 0.5, fill = c ("#798E87", "#C27D38",
     "#972D15", "#29211F")) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "n") 
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.edge", ".pdf")
ggsave(filename = file_out, plot = KM.edge, path = dir_out, device = "pdf", 
          width = 170, height = 200, units = "mm")


# Pradnik scraper edge retouch 
# Load data sheet Pradnik scraper edge retouch 
PS_edge <- read.xlsx(xlsxFile = data_file, sheet = 25)  
# Defines only the rows with complete Keilmesser and Keilmesser tips   
PS_edge <- PS_edge[-c(3:5),]

# Barplot Pradnik scraper edge retouch  
PS.edge <- ggplot(data = PS_edge) + aes(x = retouch.type.edge, 
                                     fill = retouch.type.edge) + 
     geom_bar(stat = "count", width = 0.5, fill = c ("#798E87", "#C27D38",
     "#972D15", "#29211F")) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "n") 
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.edge", ".pdf")
ggsave(filename = file_out, plot = PS.edge, path = dir_out, device = "pdf", 
          width = 170, height = 200, units = "mm")

```


## Barplot
### Raw material 
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Keilmesser raw material
# Load data sheet Keilmesser raw material
KM_raw_material <- read.xlsx(xlsxFile = data_file, sheet = 16)  

# Barplot Keilmesser raw material  
KM.raw_material <- ggplot(data = KM_raw_material) + aes(x = raw.material, 
                                                     fill = raw.material) + 
     geom_bar(stat = "count", width = 0.22, fill = c("#D69C4E", "#046C9A")) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "n") + 
     scale_x_discrete(labels=c("Baltic flint", "silicified schist")) 
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.raw_material",
                     ".pdf")
ggsave(filename = file_out, plot = KM.raw_material, path = dir_out, 
         device = "pdf")


# Pradnik scraper raw material
# Load data sheet Pradnik scraper raw material
PS_raw_material <- read.xlsx(xlsxFile = data_file, sheet = 17)  

# Barplot Pradnik scraper raw material  
PS.raw_material <- ggplot(data = PS_raw_material) + aes(x = raw.material, 
                                                     fill = raw.material) + 
     geom_bar(stat = "count", width = 0.08, fill = c("#046C9A")) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "n") + 
     scale_x_discrete(labels=c("silicified schist")) 
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.raw_material",
                   ".pdf")
ggsave(filename = file_out, plot = PS.raw_material, path = dir_out, 
         device = "pdf")


# All tool types raw material
# Load data sheet all tool types raw material
all_raw_material <- read.xlsx(xlsxFile = data_file, sheet = 1)  

# Barplot Pradnik scraper raw material  
all.raw_material <- ggplot(data = all_raw_material) + aes(x = raw.material, 
                                                       fill = raw.material) + 
     geom_bar(stat = "count", width = 0.24, fill = c("#D69C4E", "#046C9A")) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "n") + 
     scale_x_discrete(labels=c("Baltic flint", "silicified schist")) 
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "all.raw_material",
                      ".pdf")
ggsave(filename = file_out, plot = all.raw_material, path = dir_out, 
         device = "pdf")

```


## Barplot
### Morphology back 
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Keilmesser morphology back 
# Load data sheet Keilmesser morphology back
KM_back <- read.xlsx(xlsxFile = data_file, sheet = 22)  

# Barplot Keilmesser morphology back  
KM.back_morpho <- ggplot(data = KM_back) + aes(x = morphology.back, 
                                            fill = morphology.back) + 
     geom_bar(stat = "count", width = 0.5, fill = c("#518BA0", "#497C80",
     "#D69C4E", "#729394","#B9C7AD")) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "n") + 
     scale_x_discrete(labels=c("cortex + partly retouched", "cortex/unworked",
                               "N/A", "partly retouched", "retouched")) 
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.back_morpho",
                   ".pdf")
ggsave(filename = file_out, plot = KM.back_morpho, path = dir_out, 
         device = "pdf",width = 250, height = 170, units = "mm")


# Pradnik scraper morphology back 
# Load data sheet Pradnik scraper morphology back
PS_back <- read.xlsx(xlsxFile = data_file, sheet = 23)  

# Barplot Pradnik scraper morphology back 
PS.back_morpho <- ggplot(data = PS_back) + aes(x = morphology.back, 
                                            fill = morphology.back) + 
     geom_bar(stat = "count", width = 0.4, fill = c("#518BA0", "#497C80",
                                                    "#D69C4E", "#729394")) +
     theme_classic() +
     theme(legend.position = "none") + 
     labs(x = " ", y = "n") + 
     scale_x_discrete(labels=c("cortex + partly retouched", "cortex/unworked",
                               "N/A", "partly retouched", "retouched")) +
     scale_fill_manual(values = wes_palette(n = 7, name = "Darjeeling2", 
                                         type = "continuous"))    
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.back_morpho",
                   ".pdf")
ggsave(filename = file_out, plot = PS.back_morpho, path = dir_out, 
         device = "pdf",width = 250, height = 170, units = "mm")


# Keilmesser blanks
# Load data sheet Keilmesser blanks 

KM_cortex_blanks <- read.xlsx(xlsxFile = data_file, sheet = 20)  
# Ascribes the N value 
n <- doBy::summaryBy(blank ~ cortex, data = KM_cortex_blanks, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], " (n = ", n[[2]],
                                                     ")"))

# Barplot Keilmesser blanks 
KM.cortex_blanks <- ggplot(data = KM_cortex_blanks) + aes(x = blank, 
                                                       fill = cortex) + 
     geom_bar(stat = "count", width = 0.4) +
     theme_classic() +
     labs(x = " ", y = "n") + 
     scale_x_discrete(labels=c("core", "flake", "N/A")) +
     scale_fill_manual(values = wes_palette(n = 7, name = "Darjeeling2", 
                                         type = "continuous"), labels = tag)    
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), 
                   "KM.cortex_blanks", ".pdf")
ggsave(filename = file_out, plot = KM.cortex_blanks, path = dir_out, 
         device = "pdf")


# Pradnik scraper blanks
# Load data sheet Pradnik scraper blanks 
PS_cortex_blanks <- read.xlsx(xlsxFile = data_file, sheet = 21)  

# Ascribes the N value 
n <- doBy::summaryBy(blank ~ cortex, data = PS_cortex_blanks, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], 
                                                     " (n = ", n[[2]], ")"))

# Barplot Pradnik scraper blanks 
PS.cortex_blanks <- ggplot(data = PS_cortex_blanks) + aes(x = blank, 
                                                       fill = cortex) + 
     geom_bar(stat = "count", width = 0.3) +
     theme_classic() +
     labs(x = " ", y = "n") + 
     scale_x_discrete(labels=c("core", "flake", "N/A")) +
     scale_fill_manual(values = wes_palette(n = 7, name = "Darjeeling2", 
                                         type = "continuous"), labels = tag)    
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.cortex_blanks", ".pdf")
ggsave(filename = file_out, plot = PS.cortex_blanks, path = dir_out, device = "pdf")

```



## Barplot
### Pradnik method
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Keilmesser application Pradnik method
# Load data sheet Keilmesser Pradnik method
KM_Pradnik.method <- read.xlsx(xlsxFile = data_file, sheet = 30)  

# Barplot Keilmesser Pradnik method 
KM.PM <- ggplot(data = KM_Pradnik.method) + aes(x = application.Pradnik.method,
                                             fill = application.Pradnik.method) + 
     geom_bar(stat = "count", width = 0.3) + 
     theme_classic() +
     labs(x = " ", y = "n") + 
     theme(legend.position = "none") + 
     scale_x_discrete(labels=c("no", "N/A", "yes")) +
     scale_fill_manual(values = wes_palette(n = 5, name = "GrandBudapest1", 
                                         type = "continuous"))    
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.PM", ".pdf")
ggsave(filename = file_out, plot = KM.PM, path = dir_out, device = "pdf")

```


## Barplot
### Lateralisation 
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Keilmesser lateralisation
# Load data sheet Keilmesser lateralisation
KM_lateralisation <- read.xlsx(xlsxFile = data_file, sheet = 32)  
 # Barplot Keilmesser lateralisation
KM.lat <- ggplot(data = KM_lateralisation) + aes(x = tool.lateralisation, 
          fill = tool.lateralisation) + 
          geom_bar(stat = "count", width = 0.3) + 
          theme_classic() +
          labs(x = " ", y = "n") + 
          theme(legend.position = "none") +
          scale_fill_manual(values = wes_palette(n = 9, name = "Royal1", 
                                              type = "continuous"))    
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.lat", ".pdf")
ggsave(filename = file_out, plot = KM.lat, path = dir_out, device = "pdf")


# Pradnik scraper lateralisation
# Load data sheet Pradnik scraper lateralisation
PS_lateralisation <- read.xlsx(xlsxFile = data_file, sheet = 33)  

# Pradnik scraper lateralisation
# Load data sheet Pradnik scraper lateralisation
PS_lateralisation <- read.xlsx(xlsxFile = data_file, sheet = 33)
# Barplot Keilmesser lateralisation
PS.lat <- ggplot(data = PS_lateralisation) + aes(x = tool.lateralisation, 
          fill = tool.lateralisation) + 
          geom_bar(stat = "count", width = 0.3) + 
          theme_classic() +
          labs(x = " ", y = "n") + 
          theme(legend.position = "none") +
          scale_fill_manual(values = wes_palette(n = 9, name = "Royal1", 
                                              type = "continuous")) 

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "PS.lat", ".pdf")
ggsave(filename = file_out, plot = PS.lat, path = dir_out, device = "pdf")

```

## Barplot
### Barplot lateral resharpening spall type
```{r, fig.width = 10, fig.height = 8, dpi = 200}
# Lateral resharpening spall type
# Load data sheet lateral resharpening spall type
LSS_type <- read.xlsx(xlsxFile = data_file, sheet = 34) 

# Ascribes the N value 
n <- doBy::summaryBy(type.lateral.sharpening.spall ~ tool.lateralisation, 
                     data = LSS_type, FUN = length)
tag <- gsub(pattern = "_", replacement = " ", paste0(n[[1]], 
                                                     " (n = ", n[[2]], ")"))

# Barplot lateral resharpening spall type
LSS.type <- ggplot(data = LSS_type) + aes(x = type.lateral.sharpening.spall, 
                                       fill = tool.lateralisation) + 
     geom_bar(stat = "count", width = 0.2) + 
     theme_classic() +
     labs(x = " ", y = "n") + 
     labs(fill = "tool lateralisation") +
     scale_x_discrete() +
     scale_fill_manual(values = wes_palette(n = 9, name = "Royal1", 
     type = "continuous"), labels = tag)
 
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "LSS.type", ".pdf")
ggsave(filename = file_out, plot = LSS.type, path = dir_out, device = "pdf")

```



## Ternary plot
### Perimeter
```{r, fig.width = 10, fig.height = 8, dpi = 200}
library(ggtern)
# Perimeter Keilmesser 
# Load data sheet Keilmesser perimeter
KM_perimeter <- read.xlsx(xlsxFile = data_file, sheet = 10) 
# Defines only the rows with complete Keilmesser  
KM_perimeter <- KM_perimeter[1:158,]

# Ternary diagram Keilmesser perimeter
KM.perimeter  <- ggtern(data = KM_perimeter, aes(x = perimeter.distal.posterior.part, y = perimeter.active.edge, z = perimeter.basis.back)) +
                 geom_point(aes(colour = morpho.type)) +
                 theme_bw() +
                 scale_colour_startrek() +
                 theme_hidetitles() +
                 theme_showarrows() +
                 xlab("distal posterior part")+ 
                 ylab("active edge")+
                 zlab("basis + back")+ 
                 labs(colour = "Keilmesser shape") +
                 tern_limits(labels=c(0, 20, 40, 60, 80, 100)) +
                 theme_rotate(degrees = 330)
                

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "KM.perimeter", ".pdf")
ggsave(filename = file_out, plot = KM.perimeter, path = dir_out, device = "pdf")

```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("analysis/Balve/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT







