---
title: 'Plots - Lithic analysis Balver Höhle '
author: "Lisa Schunk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", 
  knit_root_dir = rprojroot::find_rstudio_root_file()) })
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script reads the xlsx file (derived data) containing all the information gained through a lithic analysis.   
The script will:

1. Reads the xlsx file 
2. Plots all relevant variables in various combinations 
3. Saves the plot as PDFs


```{r}
dir_in <-  "analysis/Balve/derived_data/"
dir_out <- "analysis/Balve/plots/"
```

Raw data must be located in "`r dir_in`".  
Formatted data will be saved in "`r dir_out`".
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(openxlsx)
library(R.utils)
library(tools)
library(chron)
library(ggplot2)
```


---

# Get name, path and information of the file
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                      row.names = NULL)
```

The checksum (MD5 hashes) of the imported files are:  
```{r, echo = FALSE}
info_in
```


# Load data into R object
```{r}
imp_data <- loadObject(paste0(dir_in, "Balve_lithic_analysis.Rbin"))
```


# Compute summary statistics
## Create function to compute the statistics at once
```{r}
nminmaxmeanmedsd <- function(x){
	y <- x[!is.na(x)]
	n_test <- length(y)
	min_test <- min(y)
	max_test <- max(y)
	mean_test <- mean(y)
 	med_test <- median(y)
 	sd_test <- sd(y)
 	out <- c(n_test, min_test, max_test, mean_test, med_test, sd_test)
 	names(out) <- c("n", "min", "max", "mean", "median", "sd")
 	return(out)
}
```

# Data analsysis - plots  
## Dimension  
### Dimesnion length

```{r}
# Define grouping variable and compute the summary statistics 
KM_length <- summaryBy(.~ ID+artefact.state+length, 
                  data=imp_data, 
                  FUN=nminmaxmeanmedsd)
str(KM_length)
```





```{r}
# histogram Keilmesser length
KM.length <- ggplot(KM_dimensions, aes(x = length, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(20, 160) +
  labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(length)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Zissou1"), labels = c("complete", "Keilmesser point", "semifinished product")) 
 
ggsave("KM.length.png") 
  
# histogram scraper length
S.length <- ggplot(S_dimensions, aes(x = length, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(40, 100) +
  labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(length)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "Zissou1"))

ggsave("S.length.png")

# histogram Pradnick scraper length
PS.length <- ggplot(PS_dimensions, aes(x = length, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(30, 80) +
  labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(length)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "FantasticFox1")) 

ggsave("PS.length.png")

# histogram lateral sharpening spall length
LSS.length <- ggplot(LSS_dimensions, aes(x = length, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(5, 50) +
  labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(length)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Darjeeling2"), labels = c("complete", "distal fragment", "medial fragment")) 

ggsave("LSS.length.png")




```

### Dimension width 
```{r}
# histogram Keilmesser width
KM.width <- ggplot(KM_dimensions, aes(x = width, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(20, 90) +
  labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(width)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Zissou1"), labels = c("complete", "Keilmesser point", "semifinished product")) 
print(KM.width) 
ggsave("KM.width.png") 
  
# histogram scraper width
S.width <- ggplot(S_dimensions, aes(x = width, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(25, 50) +
  labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(width)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "Zissou1"))

ggsave("S.width.png")

# histogram Pradnick scraper width
PS.width <- ggplot(PS_dimensions, aes(x = width, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(20, 50) +
  labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(width)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "FantasticFox1")) 

ggsave("PS.width.png")

# histogram lateral sharpening spall width
LSS.width <- ggplot(LSS_dimensions, aes(x = width, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(10, 30) +
  labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(width)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Darjeeling2"), labels = c("complete", "distal fragment", "medial fragment")) 

ggsave("LSS.width.png")


```


### Dimension thickness 
```{r}
# histogram Keilmesser thickness 
KM.thickness  <- ggplot(KM_dimensions, aes(y = thickness, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(5, 30) +
  labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_hline(aes(yintercept=mean(thickness)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Zissou1"), labels = c("complete", "Keilmesser point", "semifinished product")) 

ggsave("KM.thickness.png") 
  
# histogram scraper thickness 
S.thickness <- ggplot(S_dimensions, aes(y = thickness, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(5, 25) +
  labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_hline(aes(yintercept=mean(thickness)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "Zissou1"))

ggsave("S.thickness.png")

# histogram Pradnick scraper thickness 
PS.thickness <- ggplot(PS_dimensions, aes(y = thickness, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(0, 25) +
  labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_hline(aes(yintercept=mean(thickness)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1 ,name = "FantasticFox1")) 

ggsave("PS.thickness.png")

# histogram lateral sharpening spall thickness 
LSS.thickness <- ggplot(LSS_dimensions, aes(y = thickness, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(0, 15) +
  labs(y = "thickness[mm]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() + 
  geom_hline(aes(yintercept=mean(thickness)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Darjeeling2"), labels = c("complete", "distal fragment", "medial fragment")) 

ggsave("LSS.thickness.png")


```



### Length-width ratio 
```{r}
# scatterplott Keilmesser (complete + point) length VS width 
KM.length_width <- ggplot(KM.complete_dimensions, aes(y = length, x = width, fill = artefact.state,)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "", fill = "artefact state", size = 12) +
   xlim(0, 150) + ylim(0, 150) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =3,name = "Zissou1", type = "continuous"),labels = c("complete", "Keilmesser point")) 

ggsave("KM.length_width.png")


# scatterplott Pradnick scraper length VS width 
PS.length_width <- ggplot(PS_dimensions, aes(y = length, x = width, fill = artefact.state)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "artefact state", size = 12) +
   xlim(10, 70) + ylim(30, 80) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =1,name = "FantasticFox1", type = "continuous")) 


ggsave("PS.length_width.png")

# scatterplott lateral sharpening spall length VS width 
LSS.complete_dimensions <- LSS_dimensions[1:110,] 

LSS.length_width <- ggplot(LSS.complete_dimensions, aes(y = length, x = width, fill = artefact.state)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "artefact state", size = 12) +
   xlim(0, 50) + ylim(0, 70) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =1,name = "GrandBudapest1", type = "continuous")) 

ggsave("LSS.length_width.png")

# scatterplot Keilmesser (complete) + Pradnick scraper length VS width
KM.PS_dimensions <- dimensions[c(1:191, 309:335), ] %>% arrange(artefact.state)
KM.PS_dimensions <- KM.PS_dimensions[1:185,]

KM.PS.length_width <- ggplot(KM.PS_dimensions, aes(y = length, x = width, fill = technological.class)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "artefact type", size = 12) +
   xlim(0, 150) + ylim(0, 150) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =3,name = "GrandBudapest1", type = "continuous"), labels = c("Keilmesser", "Pradnick scraper")) 

print(KM.PS.length_width)
ggsave("KM.PS.length.width.png")


```


##

## Weight 
```{r}
# histogram Keilmesser weight
KM.weight <- ggplot(KM_weight, aes(y = weight, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(0, 1500) +
  labs(y = "weight [kg]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_hline(aes(yintercept=mean(weight)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Zissou1"), labels = c("complete", "Keilmesser point", "semifinished product")) 

print(KM.weight)



KM_weight 
PS_weight 
LSS_weight 
S_weight 


```




#### 6.4
#### 6.5
#### 6.6
#### 6.7


## Morpho type  
```{r}
# scatterplot Keilmesser (complete) + morpho types length VS width
KM.morpho.length_width <-  ggplot(KM_morpho.type, aes(y = length, x = width, fill = morpho.type)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "morpho type", size = 12) +
   xlim(0, 150) + ylim(0, 160) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =7,name = "FantasticFox1", type = "continuous"), labels = c("Bockstein", "Balve", "Klausennische","Königsaue", "Pradnick", "Buhlen", "undeterminable")) 

ggsave("KM.morpho.length_width.png")


# scatterplot Pradnick scraper (complete) + morpho types length VS width
PS.morpho.length_width <-  ggplot(PS_morpho.type, aes(y = length, x = width, fill = morpho.type)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "morpho type", size = 12) +
   xlim(0, 80) + ylim(20, 80) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =7,name = "FantasticFox1", type = "continuous"), labels = c("Bockstein", "Balve", "Klausennische","Königsaue", "Pradnick", "Buhlen", "undeterminable")) 

 
ggsave("PS.morpho.length_width.png")



```
