---
title: 'analysis techno-typological data - Balver Höhle '
author: "Lisa Schunk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", 
  knit_root_dir = rprojroot::find_rstudio_root_file()) })
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script reads the xlsx file (database techno-typological analysis) generated with E4 and formats the data for a statistical analysis.  
The script will:

1. Reads in the original xlsx file 
2. Changes the data (step #4) in order to do stats 
3. Plots the data and saves the plots as png files 


```{r}
dir_in <- "E:/EXPERIMENTS/tool-function_experiment/DATA/Inotec/R/raw_data/data_txt"
dir_out <- "analysis/derived_data/"
```

Raw data must be located in "`r dir_in`".  
Formatted data will be saved in "`r dir_out`".
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(openxlsx)
library(readxl)
library(R.utils)
library(tools)
library(data.table)
#library(plyr)
#library(reshape2)
#library(ggpubr)
#library(magrittr)
library(tidyverse)
library(wesanderson)
```


---

# Get name, path and information of the file
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                      row.names = NULL)
```

The checksum (MD5 hashes) of the imported files are:  
```{r, echo = FALSE}
info_in
```


# Read in original xlsx-file
```{r}
imp_data <- read.xlsx(xlsxFile = data_file, sheet = 1, startRow = 1, colNames = TRUE, rowNames = FALSE, skipEmptyRows = FALSE) 
#select the colums to keep 
keep_col <- c(1:2, 9:19, 21:35, 37:38, 40:43)
data_keep <- imp_data[,keep_col]

data_keep$length <- as.numeric(data_keep$length)
data_keep$width <- as.numeric(data_keep$width)
data_keep$thickness <- as.numeric(data_keep$thickness)
data_keep$weight <- as.numeric(data_keep$weight)
data_keep$perimeter.basis.back <- as.numeric(data_keep$perimeter.basis.back)
data_keep$perimeter.arch <- as.numeric(data_keep$perimeter.arch)
data_keep$perimeter.active.edge <- as.numeric(data_keep$perimeter.active.edge)
data_keep$perimeter.total <- as.numeric(data_keep$perimeter.total)
data_keep$thickness.back <- as.numeric(data_keep$thickness.back)

str(data_keep)

```

# Data analsysis - sorting  
## Dimension
```{r}
# keeps only colums relevant for dimensions and sorts them based on their technological class
keep_col <- c(2, 4:5,19:21)
dimensions <- data_keep[, keep_col] %>% arrange(technological.class)

KM_dimensions <- dimensions[1:191, ] %>% arrange(artefact.state)
KM.point_dimensions <- KM_dimensions[159:179,]
KM.only_dimensions <- KM_dimensions[1:158,]
KM.complete_dimensions <- KM_dimensions[1:179,]
PS_dimensions <- dimensions[309:335,] %>% arrange(artefact.state)
LSS_dimensions <- dimensions[192:308,] %>% arrange(artefact.state)
S_dimensions <- dimensions[336:347,] %>% arrange(artefact.state)


```

## Perimeter
```{r}
# keeps only colums relevant for perimeter measurements and sorts them based on their technological class
keep_col <- c(2, 4:5,7,23:26)
perimeter <- data_keep[, keep_col] %>% arrange(technological.class)

KM_perimeter <- perimeter[1:191, ] %>% arrange(artefact.state)
PS_perimeter <- perimeter[309:335,] %>% arrange(artefact.state)

```

## Weight
```{r}
# keeps only colums relevant for weight measurements and sorts them based on their technological class
keep_col <- c(2, 4:5, 22)
weight <- data_keep[, keep_col] %>% arrange(technological.class)

KM_weight <- weight[1:191, ] %>% arrange(artefact.state)
PS_weight <- weight[309:335,] %>% arrange(artefact.state)
LSS_weight <- weight[192:308,] %>% arrange(artefact.state)
S_weight <- weight[336:347,] %>% arrange(artefact.state)


```

## Raw material
```{r}
# keeps only colums relevant for raw material clasification and sorts them based on their technological class
keep_col <- c(2, 4:5, 3)
raw_material <- data_keep[, keep_col] %>% arrange(technological.class)

KM_raw_material <- raw_material[1:191, ] %>% arrange(artefact.state)
PS_raw_material <- raw_material[309:335,] %>% arrange(artefact.state)
LSS_raw_material <- raw_material[192:308,] %>% arrange(artefact.state)
S_raw_material <- raw_material[336:347,] %>% arrange(artefact.state)


```

## Cortex + blanks 
```{r}
# keeps only colums relevant for cortex and blank clasification and sorts them based on their technological class
keep_col <- c(2, 4:5, 6, 8:10)
cortex_blanks <- data_keep[, keep_col] %>% arrange(technological.class)

KM_cortex_blanks <- cortex_blanks[1:191, ] %>% arrange(artefact.state)
PS_cortex_blanks <- cortex_blanks[309:335,] %>% arrange(artefact.state)


```

## Back 
```{r}
# keeps only colums relevant for back modifications and sorts them based on their technological class
keep_col <- c(2, 4:5, 11)
back <- data_keep[, keep_col] %>% arrange(technological.class)

KM_back <- back[1:191, ] %>% arrange(artefact.state)
PS_back <- back[309:335,] %>% arrange(artefact.state)


```

## Edge retouch
```{r}
# keeps only colums relevant for edge retouch clasification and sorts them based on their technological class
keep_col <- c(2, 4:5, 12:13, 29)
edge_retouch <- data_keep[, keep_col] %>% arrange(technological.class)

KM_edge_retouch <- edge_retouch[1:191, ] %>% arrange(artefact.state)
PS_edge_retouch <- edge_retouch[309:335,] %>% arrange(artefact.state)
LSS_edge_retouch <- edge_retouch[192:308,] %>% arrange(artefact.state)
S_edge_retouch <- edge_retouch[336:347,] %>% arrange(artefact.state)


```

## Morpho type  
```{r}
# keeps only colums relevant for morpho type clasification and sorts them based on their technological class
keep_col <- c(2, 4:5, 7, 19:21)
morpho.type <- data_keep[, keep_col] %>% arrange(technological.class)

KM_morpho.type <- morpho.type[1:191, ] %>% arrange(artefact.state)
PS_morpho.type <- morpho.type[309:335,] %>% arrange(artefact.state)


```

## Application 'Pradnick method'  
```{r}
# keeps only colums relevant for 'morpho type'Pradnick method' clasification and sorts them based on their technological class
keep_col <- c(2, 4:5, 15:16)
Pradnick.method <- data_keep[, keep_col] %>% arrange(technological.class)

KM_Pradnick.method <- Pradnick.method[1:191, ] %>% arrange(artefact.state)
PS_Pradnick.method <- Pradnick.method[309:335,] %>% arrange(artefact.state)


```

## Lateralisation
```{r}
# keeps only colums relevant for lateralisation and sorts them based on their technological class
keep_col <- c(2, 4:5, 18)
lateralisation <- data_keep[, keep_col] %>% arrange(technological.class)

KM_lateralisation <- lateralisation[1:191, ] %>% arrange(artefact.state)
PS_lateralisation <- lateralisation[309:335,] %>% arrange(artefact.state)


```

## Type lateral sharpening spall 
```{r}
# keeps only colums relevant for lateral sharpening spall clasification and sorts them based on their technological class
keep_col <- c(2, 4:5, 17:18)
lss_type <- data_keep[, keep_col] %>% arrange(technological.class)

LSS_lss_type <- lss_type[192:308, ] %>% arrange(artefact.state)


```

# Data analsysis - stats  
## Dimension
```{r}
# descriptive statistics 
# Keilmesser (including semi-finished products, and Km-points)
length(KM_dimensions[["technological.class"]])
summary(KM_dimensions[["length"]])
summary(KM_dimensions[["width"]])
summary(KM_dimensions[["thickness"]])

# Keilmesser (complete)
length(KM.only_dimensions[["technological.class"]])
summary(KM.only_dimensions[["length"]])
summary(KM.only_dimensions[["width"]])
summary(KM.only_dimensions[["thickness"]])

# Keilmesser-points
length(KM.point_dimensions[["technological.class"]])
summary(KM.point_dimensions[["length"]])
summary(KM.point_dimensions[["width"]])
summary(KM.point_dimensions[["thickness"]])

# Pradnick scraper
length(PS_dimensions[["technological.class"]])
summary(PS_dimensions[["length"]])
summary(PS_dimensions[["width"]])
summary(PS_dimensions[["thickness"]])

# Later sharpening spall
length(LSS_dimensions[["technological.class"]])
summary(LSS_dimensions[["length"]])
summary(LSS_dimensions[["width"]])
summary(LSS_dimensions[["thickness"]])

# scraper 
length(S_dimensions[["technological.class"]])
summary(S_dimensions[["length"]])
summary(S_dimensions[["width"]])
summary(S_dimensions[["thickness"]]) 


```


# Data analsysis - plots  

## Dimension  
### Dimesnion length
```{r}
# histogram Keilmesser length
KM.length <- ggplot(KM_dimensions, aes(x = length, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(20, 160) +
  labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(length)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Zissou1"), labels = c("complete", "Keilmesser point", "semifinished product")) 
 
ggsave("KM.length.png") 
  
# histogram scraper length
S.length <- ggplot(S_dimensions, aes(x = length, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(40, 100) +
  labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(length)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "Zissou1"))

ggsave("S.length.png")

# histogram Pradnick scraper length
PS.length <- ggplot(PS_dimensions, aes(x = length, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(30, 80) +
  labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(length)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "FantasticFox1")) 

ggsave("PS.length.png")

# histogram lateral sharpening spall length
LSS.length <- ggplot(LSS_dimensions, aes(x = length, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(5, 50) +
  labs(x = "length [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(length)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Darjeeling2"), labels = c("complete", "distal fragment", "medial fragment")) 

ggsave("LSS.length.png")




```

### Dimension width 
```{r}
# histogram Keilmesser width
KM.width <- ggplot(KM_dimensions, aes(x = width, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(20, 90) +
  labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(width)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Zissou1"), labels = c("complete", "Keilmesser point", "semifinished product")) 
print(KM.width) 
ggsave("KM.width.png") 
  
# histogram scraper width
S.width <- ggplot(S_dimensions, aes(x = width, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(25, 50) +
  labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(width)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "Zissou1"))

ggsave("S.width.png")

# histogram Pradnick scraper width
PS.width <- ggplot(PS_dimensions, aes(x = width, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(20, 50) +
  labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(width)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "FantasticFox1")) 

ggsave("PS.width.png")

# histogram lateral sharpening spall width
LSS.width <- ggplot(LSS_dimensions, aes(x = width, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + xlim(10, 30) +
  labs(x = "width [mm]", y = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_vline(aes(xintercept=mean(width)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Darjeeling2"), labels = c("complete", "distal fragment", "medial fragment")) 

ggsave("LSS.width.png")


```


### Dimension thickness 
```{r}
# histogram Keilmesser thickness 
KM.thickness  <- ggplot(KM_dimensions, aes(y = thickness, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(5, 30) +
  labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_hline(aes(yintercept=mean(thickness)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Zissou1"), labels = c("complete", "Keilmesser point", "semifinished product")) 

ggsave("KM.thickness.png") 
  
# histogram scraper thickness 
S.thickness <- ggplot(S_dimensions, aes(y = thickness, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(5, 25) +
  labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_hline(aes(yintercept=mean(thickness)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1,name = "Zissou1"))

ggsave("S.thickness.png")

# histogram Pradnick scraper thickness 
PS.thickness <- ggplot(PS_dimensions, aes(y = thickness, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(0, 25) +
  labs(y = "thickness [mm]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_hline(aes(yintercept=mean(thickness)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =1 ,name = "FantasticFox1")) 

ggsave("PS.thickness.png")

# histogram lateral sharpening spall thickness 
LSS.thickness <- ggplot(LSS_dimensions, aes(y = thickness, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(0, 15) +
  labs(y = "thickness[mm]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() + 
  geom_hline(aes(yintercept=mean(thickness)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Darjeeling2"), labels = c("complete", "distal fragment", "medial fragment")) 

ggsave("LSS.thickness.png")


```



### Length-width ratio 
```{r}
# scatterplott Keilmesser (complete + point) length VS width 
KM.length_width <- ggplot(KM.complete_dimensions, aes(y = length, x = width, fill = artefact.state,)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "", fill = "artefact state", size = 12) +
   xlim(0, 150) + ylim(0, 150) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =3,name = "Zissou1", type = "continuous"),labels = c("complete", "Keilmesser point")) 

ggsave("KM.length_width.png")


# scatterplott Pradnick scraper length VS width 
PS.length_width <- ggplot(PS_dimensions, aes(y = length, x = width, fill = artefact.state)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "artefact state", size = 12) +
   xlim(10, 70) + ylim(30, 80) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =1,name = "FantasticFox1", type = "continuous")) 


ggsave("PS.length_width.png")

# scatterplott lateral sharpening spall length VS width 
LSS.complete_dimensions <- LSS_dimensions[1:110,] 

LSS.length_width <- ggplot(LSS.complete_dimensions, aes(y = length, x = width, fill = artefact.state)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "artefact state", size = 12) +
   xlim(0, 50) + ylim(0, 70) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =1,name = "GrandBudapest1", type = "continuous")) 

ggsave("LSS.length_width.png")

# scatterplot Keilmesser (complete) + Pradnick scraper length VS width
KM.PS_dimensions <- dimensions[c(1:191, 309:335), ] %>% arrange(artefact.state)
KM.PS_dimensions <- KM.PS_dimensions[1:185,]

KM.PS.length_width <- ggplot(KM.PS_dimensions, aes(y = length, x = width, fill = technological.class)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "artefact type", size = 12) +
   xlim(0, 150) + ylim(0, 150) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =3,name = "GrandBudapest1", type = "continuous"), labels = c("Keilmesser", "Pradnick scraper")) 

print(KM.PS.length_width)
ggsave("KM.PS.length.width.png")


```


##

## Weight 
```{r}
# histogram Keilmesser weight
KM.weight <- ggplot(KM_weight, aes(y = weight, fill = artefact.state)) + 
  geom_histogram(binwidth = 1) + ylim(0, 1500) +
  labs(y = "weight [kg]", x = "N", title = "", fill = "artefact state", size = 12) +
  theme_classic() +
  geom_hline(aes(yintercept=mean(weight)), linetype="dashed", size=1) +
  scale_fill_manual(values = wes_palette(n =3,name = "Zissou1"), labels = c("complete", "Keilmesser point", "semifinished product")) 

print(KM.weight)



KM_weight 
PS_weight 
LSS_weight 
S_weight 


```




#### 6.4
#### 6.5
#### 6.6
#### 6.7


## Morpho type  
```{r}
# scatterplot Keilmesser (complete) + morpho types length VS width
KM.morpho.length_width <-  ggplot(KM_morpho.type, aes(y = length, x = width, fill = morpho.type)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "morpho type", size = 12) +
   xlim(0, 150) + ylim(0, 160) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =7,name = "FantasticFox1", type = "continuous"), labels = c("Bockstein", "Balve", "Klausennische","Königsaue", "Pradnick", "Buhlen", "undeterminable")) 

ggsave("KM.morpho.length_width.png")


# scatterplot Pradnick scraper (complete) + morpho types length VS width
PS.morpho.length_width <-  ggplot(PS_morpho.type, aes(y = length, x = width, fill = morpho.type)) +
   geom_point(size = 3, shape = 21) +
   labs(x = "length [mm]", y = "width [mm]", title = "",fill = "morpho type", size = 12) +
   xlim(0, 80) + ylim(20, 80) +
   theme_classic() +
   scale_fill_manual(values = wes_palette(n =7,name = "FantasticFox1", type = "continuous"), labels = c("Bockstein", "Balve", "Klausennische","Königsaue", "Pradnick", "Buhlen", "undeterminable")) 

 
ggsave("PS.morpho.length_width.png")



```




# Save data
## Define output file name
```{r}
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_IE_plot", ".pdf")
ggsave(filename = file_out, plot = data_plot, path = dir_out, device = "pdf")

```
The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".

---


# Show plot files information
```{r}
info_out <- list.files(path = dir_out, pattern = "\\.pdf$", 
                       full.names = TRUE) %>% 
            md5sum()
```


The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```

---

# sessionInfo() and RStudio version
```{r}
sessionInfo()
```

---

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


--- 

END OF SCRIPT