---
title: 'Import - Lithic analysis Ramioul'
author: "Lisa Schunk"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)

```


---

# Goal of the script
This script reads the xlsx file (database techno-typological analysis) generated with E4
and formats the data for a statistical analysis.  
The script will:

1. Reads in the original xlsx file 
2. Changes and sort the data  in order to do stats 
3. Saves the data as a new xlsx file and R object


```{r}
dir_in <-  "analysis/Ramioul/raw_data/"
dir_out <- "analysis/Ramioul/derived_data/"

```

Raw data must be located in "`r dir_in`".  
Formatted data will be saved in "`r dir_out`".
The knit directory for this script is the project directory.

---


# Load packages
```{r Libraries}
pack_to_load <- c("R.utils", "tools", "openxlsx", "data.table", "readxl", "chron", "dplyr")
sapply(pack_to_load, library, character.only = TRUE, logical.return = TRUE)

```


---

# Get name, path and information of the file
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                      row.names = NULL)

```



# Read in original xlsx-file
```{r}
imp_data <- read.xlsx(xlsxFile = data_file, sheet = 1, startRow = 1, colNames = TRUE,
                      rowNames = FALSE, skipEmptyRows = FALSE) 
#select the columns to keep 
keep_col <- c(1:2, 9:19, 21:35, 37:38, 40:43)
data_final <- imp_data[,keep_col]

data_final$length <- as.numeric(data_final$length)
data_final$width <- as.numeric(data_final$width)
data_final$thickness <- as.numeric(data_final$thickness)
data_final$weight <- as.numeric(data_final$weight)
data_final$perimeter.basis.back <- as.numeric(data_final$perimeter.basis.back)
data_final$perimeter.arch <- as.numeric(data_final$perimeter.arch)
data_final$perimeter.active.edge <- as.numeric(data_final$perimeter.active.edge)
data_final$perimeter.total <- as.numeric(data_final$perimeter.total)
data_final$thickness.back <- as.numeric(data_final$thickness.back)

str(data_final)

```


# Data analsysis - sorting  
## Dimension
```{r}
# keep only columns relevant for dimensions and sort them based on 
# their technological class
keep_col <- c(1:2, 4:5,19:21)
dimensions <- data_final[, keep_col] %>% arrange(technological.class)

KM_dimensions <- dimensions[3:11, ] 
PS_dimensions <- dimensions[12:14, ] 
F_dimensions <- dimensions[1:2, ]
S_dimensions <- dimensions[15:20, ] 

```


## Perimeter
```{r}
# keep only columns relevant for perimeter measurements and sort them 
# based on their technological class
keep_col <- c(1:2, 4:5,7,23:26)
perimeter <- data_final[, keep_col] %>% arrange(technological.class)

KM_perimeter <- perimeter[3:11, ] 
PS_perimeter <- perimeter[12:14, ] 

```


## Weight
```{r}
# keep only columns relevant for weight measurements and sort them based 
# on their technological class
keep_col <- c(1:2, 4:5, 22)
weight <- data_final[, keep_col] %>% arrange(technological.class)

KM_weight <- weight[3:11, ] 
PS_weight <- weight[12:14, ] 
F_weight <- weight[1:2, ] 
S_weight <- weight[15:20, ] 

```


## Raw material
```{r}
# keep only columns relevant for raw material classification and sort them 
# based on their technological class
keep_col <- c(1:2, 4:5, 3)
raw_material <- data_final[, keep_col] %>% arrange(technological.class)

KM_raw_material <- raw_material[3:11, ] 
PS_raw_material <- raw_material[12:14, ]  
F_raw_material <- raw_material[1:2, ] 
S_raw_material <- raw_material[15:20, ] 

```


## Cortex + blanks 
```{r}
# keep only columns relevant for cortex and blank classification and sort them 
# based on their technological class
keep_col <- c(1:2, 4:5, 6, 8:10)
cortex_blanks <- data_final[, keep_col] %>% arrange(technological.class)

KM_cortex_blanks <- cortex_blanks[3:11, ] 
PS_cortex_blanks <- cortex_blanks[12:14, ] 

```


## Back 
```{r}
# keep only columns relevant for back modifications and sort them based 
# on their technological class
keep_col <- c(1:2, 4:5, 11, 27)
back <- data_final[, keep_col] %>% arrange(technological.class)

KM_back <- back[3:11, ]  
PS_back <- back[12:14, ]

```


## Edge retouch
```{r}
# keep only columns relevant for edge retouch classification and sort them 
# based on their technological class
keep_col <- c(1:2, 4:5, 12:13, 29)
edge_retouch <- data_final[, keep_col] %>% arrange(technological.class)

KM_edge_retouch <- edge_retouch[3:11, ]
PS_edge_retouch <- edge_retouch[12:14, ] 
F_edge_retouch <- edge_retouch[1:2, ] 
S_edge_retouch <- edge_retouch[15:20, ] 

```


## Morpho type  
```{r}
# keep only columns relevant for morpho type classification and sort them 
# based on their technological class
keep_col <- c(1:2, 4:5, 7, 19:21)
morpho.type <- data_final[, keep_col] %>% arrange(technological.class)

KM_morpho.type <- morpho.type[3:11, ] 
PS_morpho.type <- morpho.type[12:14, ] 

```


## Application 'Pradnik method'  
```{r}
# keep only columns relevant for 'morpho type 'Pradnik method' classification 
# and sort them based on their technological class
keep_col <- c(1:2, 4:5, 15:16)
Pradnik.method <- data_final[, keep_col] %>% arrange(technological.class)

KM_Pradnik.method <- Pradnik.method[3:11, ]
PS_Pradnik.method <- Pradnik.method[12:14, ]  

```


## Lateralisation
```{r}
# keep only columns relevant for lateralisation and sort them based on their
# technological class
keep_col <- c(1:2, 4:5, 18)
lateralisation <- data_final[, keep_col] %>% arrange(technological.class)

KM_lateralisation <- lateralisation[3:11, ] 
PS_lateralisation <- lateralisation[12:14, ] 

```


# Save data
## Format name of output file
```{r}
file_out <- "Ramioul_lithic_analysis"

```
The files will be saved as "`r paste0("~/", file_out, ".[ext]")`".


## Write to XLSX
```{r}
write.xlsx(list(data = data_final, dimensions = dimensions, KM_dimensions = KM_dimensions,
                PS_dimensions = PS_dimensions, F_dimensions = F_dimensions, S_dimensions =
                  S_dimensions, KM_perimeter = KM_perimeter, PS_perimeter = PS_perimeter,
                  KM_weight = KM_weight, PS_weight = PS_weight, F_weight = F_weight,
                  S_weight = S_weight, KM_raw_material = KM_raw_material,PS_raw_material =
                  PS_raw_material, F_raw_material = F_raw_material, S_raw_material =
                  S_raw_material, KM_cortex_blanks = KM_cortex_blanks, PS_cortex_blanks =
                  PS_cortex_blanks, KM_back = KM_back, PS_back = PS_back, KM_edge_retouch
                  = KM_edge_retouch, PS_edge_retouch = PS_edge_retouch, F_edge_retouch =
                  F_edge_retouch, S_edge_retouch = S_edge_retouch, KM_morpho.type =
                  KM_morpho.type, PS_morpho.type = PS_morpho.type, KM_Pradnik.method =
                  KM_Pradnik.method, PS_Pradnik.method = PS_Pradnik.method,
                  KM_lateralisation = KM_lateralisation, PS_lateralisation =
                  PS_lateralisation),
                  file = paste0(dir_out, file_out, ".xlsx"))

```


## Save R object
```{r}
saveObject(data_final, file = paste0(dir_out, file_out, ".Rbin"))

```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("RStudioVersion.txt", n = 1)`.


# Cite R packages used
```{r Citation}
for (i in pack_to_load) print(citation(i), bibtex = FALSE)

```

---


END OF SCRIPT